### 2015-07-31 tests/fitness.tex
tex

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\begin{document}
\paragraph{Assessing the quality of reconstructed functions}
In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.

A measure that is timeshift independent and maximized for functions of identical shape is the 
maximum of the cross correlation 
\begin{align}
f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
\end{align}
Algorithmically, the signals are split into chunks which are correlated separately. This saves computational power and 
prevents minute differences in frequency to accumulate into a huge phase shift, i.e. the phase only has to remain relatively
constant over the timescale of a single chunk.

\end{document}

### 2015-07-31 paper/paper.tex

\documentclass[letter,scriptaddress,twocolumn, prl,showkeys]{revtex4-1}

    \usepackage{microtype}
    \usepackage{blindtext}

	\usepackage{amsmath}%,amssymb} 
	\usepackage{makeidx}
	\usepackage{amsfonts}
	\usepackage[ansinew]{inputenc}
	\usepackage[usenames,dvipsnames]{pstricks}
	\usepackage{subfigure}
	\usepackage{epsfig}
	\usepackage{pst-grad} % For gradients
	\usepackage{pst-plot} % For axes
	\usepackage[colorlinks,hyperindex]{hyperref}
	\hypersetup
	{
		colorlinks,%
		citecolor=black,%
		linkcolor=black,%
		urlcolor=black,%
	}

%--- Theorem like environments ----
	\newtheorem{theorem}{Theorem}
	\newtheorem{corolary}{Corolary}
	\newtheorem{prop}{Proposition}
	\newtheorem{definition}{Definition}
	\newtheorem{example}{Example}
	\newtheorem{exercise}{Exercise}
	\newtheorem{lemma}{Lemma}

%--- Definindo algumas frescuras.
%\numberwithin{equation}{subsection}
	%\numberwithin{equation}{section}

	\setlength\textheight{24.5cm}

\makeindex

%--------------------------------------------------------
\begin{document}

\title{Optimizing Structural Properties Of Neural Networks With Evolutionary Algorithms}

\author{T. Staudt, E. Schultheis}
\email{thomas.staudt@stud.uni-goettingen.de, erik.schultheis@uni-goettingen.de}
\affiliation{University of Göttingen}

\date{today}

\begin{abstract}
    In this article we examine how structural properties of neural
    networks, like the size or local synapse densities, affect their
    learning success for various tasks. To do so, we look at a rate based
    model for neural networks and apply the FORCE rule for the learning
    process. A sophisticated matching algorithm
    allows us to quantify the learning success of a network so that the
    fitness of structural characteristics, expressed by integer or
    floating-point parameters, can be evaluated. We then use evolutionary
    optimization methods on these parameters to show that (1) ring
    topologies are generally more successful than strictly random
    topologies, (2) *FAT
    LIST OF INCREDIBLE RESULTS THAT CHANGE THE UNIVERSE*
\end{abstract}

\keywords{genetic algorithms, evolutionary algorithms, FORCE learning, rate networks}

\maketitle

\section{Introduction}
What makes artificial neuronal networks learn successfully? And why are
certain types of networks and certain learning rules effective for some
problems, but fail at solving others?
Confronting and eventually answering these questions is critical when trying
to gain a deep understanding of neural networks and artificial
intelligence.

*MISSING*

Instead of fine tuning single internal weights in order to obtain a network
perfectly fit for the task, however, we are concerned with the rules that
generate the structure of the networks. So we still want our networks to be
random to a certain degree, which seems biologically plausible, but we also
want them to follow certain patterns and allow specialization: How dense
are the synapses distributed? Is there a strong compartimization?  Are
recurrent structures frequent? How strong should the feedback signal be,
and how strong should the synapses be on average?

\section{Concepts and Methods}
\paragraph{Basic Network Dynamics} As our neuronal network model we choose
basic rate networks that are also used in the original FORCE publication
\cite{FORCE} and the publication \cite{FORCE}, where our notation of this
paragraph mainly stems from. So we look at

- We chose rate networks in order to apply FORCE rule
- borrow notation from reward modulation rule paper
Basic Network dynamics:
    - Introduce single neuronal network as collection of matrices with specified
    dynamics
    - Briefly describe the learning rule (FORCE)

\paragraph{Genes and Challenges}
    - Introduce parameterized random networks as generators of these networks
    - Introduce challenges, and the expected success values when paring
    generators with challenges.

\paragraph{Genetic Optimization}
Genetic Optimization:

- Hint / vague description of the codebase

\section{Results}

\section{Discussion}

\blindtext

\begin{thebibliography}{}

\bibitem{mcdermott}
  Y.-F.~Chen {\it et al.},
  %``Microwave Photon Counter Based on Josephson Junctions,''
  Phys.\ Rev.\ Lett.\  {\bf 107}, 217401 (2011).

\end{thebibliography}

\end{document}

### 2015-07-31 paper/paper.tex
introduction extended

@@
 
-\title{Optimizing Structural Properties Of Neural Networks With Evolutionary Algorithms}
+\title{Optimizing Structural Properties Of Neural Networks With Genetic Algorithms}
 
@@
 \section{Introduction}
-What makes artificial neuronal networks learn successfully? And why are
-certain types of networks and certain learning rules effective for some
-problems, but fail at solving others?
-Confronting and eventually answering these questions is critical when trying
-to gain a deep understanding of neural networks and artificial
-intelligence.
+In contemporary artificial neuronal network science a vast number of
+different models for describing both the dynamic of networks as well as
+their plasticities are applied. 
+But what exactly makes artificial neuronal networks learn successfully? And
+why are certain types of networks and certain learning rules effective for
+some problems, but fail at solving others? Confronting and eventually
+answering these questions is critical when trying to gain a deep
+understanding of neural networks and artificial intelligence.
 
-*MISSING*
+Using the above questions as a guideline,
+we decided to analyze the FORCE algorithm introduced in \cite{FORCE} that
+is especially suitable for learning periodic patterns. More precisely, we
+wanted to identify which properties of a network are important for its
+learning success, and how to choose these properties' values for an optimal
+result.  To this end, one option would be to take an initial network and
+an adequate periodic pattern to be learned, and to optimize the network's
+internals for this very task. Then patterns may emerge that hint at how an
+optimal network design looks like.
 
-Instead of fine tuning single internal weights in order to obtain a network
-perfectly fit for the task, however, we are concerned with the rules that
-generate the structure of the networks. So we still want our networks to be
-random to a certain degree, which seems biologically plausible, but we also
-want them to follow certain patterns and allow specialization: How dense
-are the synapses distributed? Is there a strong compartimization?  Are
-recurrent structures frequent? How strong should the feedback signal be,
-and how strong should the synapses be on average?
+However, we took a different approach:
+Instead of fine tuning single internal weights, we were concerned with the
+general rules of the network's structure. So we wanted our networks to
+be random to a certain degree, which seems biologically plausible, but we
+also wanted them to follow certain patterns and allow for specialization:
+How dense are the synapses distributed? Is there a strong compartmentalization?
+Are recurrent structures frequent? How strong should the feedback signal
+be, and how strong should the synapses fire on average?
+
+To address these issues we looked at \emph{parameterized random networks}, or
+in the language of genetics, at different \emph{network genotypes}. These
+genotypes contain parametric information about the networks to be created,
+the \emph{phenotypes}. Different phenotypes with the same genotype may thus
+have strong quantitative differences, but they share structural qualities
+whose fitness for a preassigned task we wanted to analyze and improve.
+
+Since calculating the fitness of a genotype required the evaluation of many
+different phenotypes and because of the resulting stochastic nature of the
+problems we faced, we decided to use genetic algorithms for the
+optimization process. Other more direct optimization methods like simulated
+annealing or gradient descent seemed unfeasible for this.
 
@@
 paragraph mainly stems from. So we look at
+
+- describe network dynamic
+- describe learning and tasks
+- describe sucess of network
 

### 2015-07-31 paper/paper.tex
test

--- 2015-07-31 paper/paper.tex
@@
 	\usepackage{epsfig}
-	\usepackage{pst-grad} % For gradients
+%	\usepackage{pst-grad} % For gradients
 	\usepackage{pst-plot} % For axes
@@
 
+\paragraph{Tasks}
+For the optimization to work properly, the problems presented to the networks 
+have (for the most part) to solvable by FORCE learning. While it was shown \cite{}
+that large recurrent neural networks are capable of learning very complex functions
+using the FORCE algorithm, the optimization requires a large number of learnings.
+Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
+and as such are less powerful.
+
+To get a simple criterion to estimate whether it might be possible to learn a given function 
+using FORCE with 100 neurons, we look at
+
 \section{Results}

### 2015-07-31 tests/understand_quality.jl
testfile commited

@@ -1,20 +1,26 @@
-include("../src/EvoNet.jl")
+#=include("../src/EvoNet.jl")=#
+using EvoNet
 ev = EvoNet
-import EvoNet.@rec
 
-N = 10
+N = 100
 
-f(t) = cos(t) + 0.5sin(3t)
-g(t) = 0
-task = ev.FunctionTask( [f;], [g;] )
+ch = ev.simple_wave(n=1, amplitude=1, frequency=0.314/2pi, offset=0, phase=0)
+task = ev.get_task(ch)
+
 
 quals = Float64[]
-rule  = ev.ForceRule( N, 1 )
+rule  = ev.Learning.ForceRule( N, α=1/100 )
 
-generator = ev.SparseMatrixGenerator( N, 0.1, gain = 1.1  )
-for i in 1:10
-    net = ev.generate( generator, i )
-    push!( quals, ev.test_fitness_for_task(net, rule, task, 
-           learntime = 500, evaltime = 1000, waittime =  200, α=1/1000,
-           fname="")[1] )
-end
+ertop = ev.ErdösRenyiTopology(0.1)
+generator = ev.SparseFRGenerator( N, topology=ertop )
+
+
+net = ev.generate( generator, seed=ev.randseed() )
+ev.test_fitness_for_task(net, rule, task, fname="simu.dat")
+
+#=for i in 1:10=#
+    #=net = ev.generate( generator, i )=#
+    #=push!( quals, ev.test_fitness_for_task(net, rule, task, =#
+           #=learntime = 500, evaltime = 1000, waittime =  200, α=1/1000,=#
+           #=fname="")[1] )=#
+#=end=#

### 2015-07-31 src/learning/evaluator.jl
zeug

@@ -53,7 +53,7 @@ end
 function calculate_correlation!( evl::Evaluator )
   @assert evl.T == evl.chunksize + 1 "trying to calculate correlation before chunk was filled"
   # now evaluate the collected chunk.
-  # first, we remove a costant phase shift. to do that, 
+  # first, we remove a costant phase shift. to do that,
   # we look at the crosscorrelation of expected and received
   # and choose Δt such that it is maximized
   mxsum = -Inf
@@ -97,8 +97,9 @@ function evaluate(evl::Evaluator, task::AbstractTask, duration::Real; rec::Bool=
     evaluate_step!(evl, task)
     if rec
       record(recorder, 1, evl.net.time)
+      record(recorder, 2, evl.net.time - evl.timeshift)
       for i = 1:length(evl.net.output)
-        record(recorder, 1+i, evl.net.output[i])
+        record(recorder, 2+i, evl.net.output[i])
       end
     end
   end

### 2015-07-31 src/fitness.jl
record estimated time

@@ -48,8 +48,9 @@ function test_fitness_for_task( net::AbstractNetwork, rule::AbstractRule,
     # record in case that a filename was given
     if fname != ""
         record(recorder, 1, net.time)
+        record(recorder, 2, net.time)
         for i = 1:length(net.output)
-           record(recorder, 1+i, net.output[1])
+           record(recorder, 2+i, net.output[1])
         end
     end
   end

### 2015-07-31 src/fitness.jl
record expected

@@ -49,8 +49,9 @@ function test_fitness_for_task( net::AbstractNetwork, rule::AbstractRule,
     if fname != ""
         record(recorder, 1, net.time)
         record(recorder, 2, net.time)
+        record(recorder, 3, get_expected(task))
         for i = 1:length(net.output)
-           record(recorder, 2+i, net.output[1])
+           record(recorder, 3+i, net.output[1])
         end
     end
   end

### 2015-07-31 src/learning/evaluator.jl

@@ -98,8 +98,9 @@ function evaluate(evl::Evaluator, task::AbstractTask, duration::Real; rec::Bool=
     if rec
       record(recorder, 1, evl.net.time)
       record(recorder, 2, evl.net.time - evl.timeshift)
+      record(recorder, 3, get_expected(task))
       for i = 1:length(evl.net.output)
-        record(recorder, 2+i, evl.net.output[i])
+        record(recorder, 3+i, evl.net.output[i])
       end
     end
   end

### 2015-07-31 src/learning/evaluator.jl
safety commit

@@ -81,6 +81,9 @@ function calculate_correlation!( evl::Evaluator )
 
   # just nice variable names
   evl.timeshift += dtime * dt
+  if mxsum / recvn / norm(evl.expected) > 0.1
+      println("$mxsum $recvn $(evl.timeshift)")
+  end
   evl.last_result = mxsum / norm(evl.expected) / recvn
   evl.sumcor += evl.last_result
   evl.chunkcount += 1
@@ -91,6 +94,7 @@ end
 
 function evaluate(evl::Evaluator, task::AbstractTask, duration::Real; rec::Bool=false, recorder=REC)
   start_time = evl.net.time
+  println(start_time)
   while evl.net.time < start_time + duration
     prepare_task!(task, evl.net.time  + dt, false) # non deterministic: allows noise in input
     update!(evl.net, get_input(task))

### 2015-07-31 src/learning/Learning.jl
get_expected

@@ -14,4 +14,4 @@ export AbstractTask, FunctionTask,
        AbstractTeacher, Teacher,
        AbstractEvaluator, Evaluator
 # functions
-export evaluate, reset, learn!
+export evaluate, reset, learn!, get_expected

### 2015-07-31 research/q_of_omega.jl
more stuff

@@ -19,6 +19,6 @@ ev = EvoNet
   return EvoNet.get_value(s)
 end
 
-freq = logspace(-2, 2, 100)
+freq = logspace(-2, 1, 100)
 result = pmap(x->rating(x), freq)
 writedlm("freqrange.txt", [freq result])

### 2015-07-31 src/learning/evaluator.jl

@@ -98,7 +98,7 @@ function evaluate(evl::Evaluator, task::AbstractTask, duration::Real; rec::Bool=
     if rec
       record(recorder, 1, evl.net.time)
       record(recorder, 2, evl.net.time - evl.timeshift)
-      record(recorder, 3, get_expected(task))
+      record(recorder, 3, get_expected(task)[1])
       for i = 1:length(evl.net.output)
         record(recorder, 3+i, evl.net.output[i])
       end

### 2015-07-31 paper/paper.tex

@@
 To get a simple criterion to estimate whether it might be possible to learn a given function 
-using FORCE with 100 neurons, we look at
+using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
+determine the range of $\omega$ in which an (unoptimized) network (i.e. GAIN, 
+FEEDBACK, p, Erdös Renyi) is able to learn the function (see fig. \ref{}).
 

### 2015-08-01 paper/paper.tex
text

@@
 
+\paragraph{Assessing the quality of reconstructed functions}
+In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
+the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
+$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
+produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.
+
+A measure that is timeshift independent and maximized for functions of identical shape is the 
+maximum of the cross correlation 
+\begin{align}
+f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
+\end{align}
+Algorithmically, the signals are split into chunks which are correlated separately. This saves computational power and 
+prevents minute differences in frequency to accumulate into a huge phase shift, i.e. the phase only has to remain relatively
+constant over the timescale of a single chunk.
+
 \section{Results}

### 2015-08-01 src/learning/evaluator.jl
more carefully defined cross correlation

@@ -30,7 +30,7 @@ end
 
 # this function does not advance network time
 # returns true is we have collected a full chunk and updated the error
-function evaluate_step!(evl::Evaluator, task::AbstractTask) # ::Bool
+function evaluate_step!(evl::Evaluator, task::AbstractTask, allow_shift::Bool=true) # ::Bool
   # update task for currently needed time
   old_time = task.time
   old_det  = task.deterministic
@@ -44,49 +44,37 @@ function evaluate_step!(evl::Evaluator, task::AbstractTask) # ::Bool
 
   evl.T += 1
   if evl.T > evl.chunksize
-    calculate_correlation!(evl)
+    calculate_correlation!(evl, allow_shift)
     return true
   end
   return false
 end
 
-function calculate_correlation!( evl::Evaluator )
+function calculate_correlation!( evl::Evaluator, allow_shift = true )
   @assert evl.T == evl.chunksize + 1 "trying to calculate correlation before chunk was filled"
   # now evaluate the collected chunk.
   # first, we remove a costant phase shift. to do that,
   # we look at the crosscorrelation of expected and received
   # and choose Δt such that it is maximized
-  mxsum = -Inf
-  dtime = 0
-  recvn = norm(evl.received)
-  # in case the network died down, we can skip all the computations (and avoid NaN)
-  if recvn == 0
-     # dead network counts as fully failed
-    evl.last_result = 0
-    evl.chunkcount += 1
-    evl.T = 1
-    return
+
+  mshift =  div(evl.chunksize, 10)
+  if !allow_shift
+    mshift = 0
   end
-  for ΔT = -div(evl.chunksize, 10):div(evl.chunksize, 10)
-    summed = 0.0
-    for i = 1+max(0,-ΔT):evl.chunksize-max(0, ΔT)
-      #println(i, i+ΔT)
-      summed += sum(evl.expected[i] .* evl.received[i + ΔT])
-    end
-    if summed > mxsum
-      mxsum = summed
-      dtime = ΔT
-    end
-  end
+
+  ΔT, corr = best_shift(evl.received, evl.expected, mshift)
 
   # just nice variable names
-  evl.timeshift += dtime * dt
-  evl.last_result = mxsum / norm(evl.expected) / recvn
-  evl.sumcor += evl.last_result
-  evl.chunkcount += 1
+  evl.timeshift   += ΔT * dt
+  evl.last_result  = corr
[…]

### 2015-08-01 paper/paper.tex
completed some text

--- 2015-07-31 paper/paper.tex
@@
 
+    \newcommand{\ix}[1]{_\mathrm{#1}}
+    \newcommand{\Ix}[1]{^\mathrm{#1}}
+
 	\setlength\textheight{24.5cm}
@@
 \section{Introduction}
-In contemporary artificial neuronal network science a vast number of
-different models for describing both the dynamic of networks as well as
-their plasticities are applied. 
+Looking at the field of artificial neuronal networks, one finds a great
+variety of different models describing both the dynamic of networks as well
+as their plasticities, i.e.\ their learning behavior.
 But what exactly makes artificial neuronal networks learn successfully? And
@@
 
-Using the above questions as a guideline,
-we decided to analyze the FORCE algorithm introduced in \cite{FORCE} that
-is especially suitable for learning periodic patterns. More precisely, we
-wanted to identify which properties of a network are important for its
-learning success, and how to choose these properties' values for an optimal
-result.  To this end, one option would be to take an initial network and
-an adequate periodic pattern to be learned, and to optimize the network's
-internals for this very task. Then patterns may emerge that hint at how an
-optimal network design looks like.
+Using the above questions as a guideline, we decided to analyze what
+network characteristics makes the FORCE algorithm, introduced in
+\citep{FORCE}, especially suitable for learning periodic patterns.  More
+precisely, we wanted to identify which properties of a network are
+important for its learning success, and how to choose these properties'
+values for an optimal result.  To this end, one option would be to take an
+initial network and an adequate periodic pattern to be learned, and to
+optimize the network's internals for this very task. Then patterns may
+emerge that hint at how an optimal network design looks like.
 
@@
 \section{Concepts and Methods}
-\paragraph{Basic Network Dynamics} As our neuronal network model we choose
-basic rate networks that are also used in the original FORCE publication
-\cite{FORCE} and the publication \cite{FORCE}, where our notation of this
-paragraph mainly stems from. So we look at
+\paragraph{Network Model} We chose the same rate based network architecture
+that is used in the original FORCE publication \citep{FORCE} (denoted as
+architecture A) and also in \citep{RM}, where our notation mainly stems
+from. So we look at a network $\mathcal{N}$ with $N$ internal neurons and
+states $x_i$ for $1 \le i \le N$ that loosely represent the neurons
+membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
+given by $r_i = \tanh x_i$.  Further, the network contains $R$ readout
+neurons with states $z_i$ for $1 \le i \le R$, where
+\begin{equation} 
+    z_i = \sum_{j=1}^{N}\omega\Ix{read}_{ij}\,r_j~.
+\end{equation}
+Taking into account both internal dynamics as well as an external feedback
+pathway, the dynamics $t \mapsto x_i(t)$ of the single neurons are governed
+by \footnote{~Note that we did not include external input signals in this
+formula, as is done in \citep{FORCE} and \citep{RM}. We in fact enabled
+inputs in our code base, but did not use them for the main results.}
+\begin{equation}
+    \dot{x}_i(t) = -x_i(t) + \sum_{i=1}^N \omega\Ix{rec}_{ij}\, r_j(t) + 
+                   \sum_{i=1}^R \omega\Ix{fb}_{ij}\, z_j(t)~,
+\end{equation}
+where we introduce the internal recurrent synapse weights
+$\omega\Ix{rec}_{ij}$ and the feedback synapse weights
+$\omega\Ix{fb}_{ij}$. One can see that the neurons exponentially
+decay if they are not presented stimulus by other neurons. 
+We solved this system of differential equations by
+a simple Newton method with integration time steps $dt = 0.01$, which was
+also used in \citep{FORCE}. (How we constructed $\omega\Ix{rec}$ and
+$\omega\Ix{fb}$ will be discussed in \emph{Parametric Random Networks})
 
-- describe network dynamic
-- describe learning and tasks
-- describe sucess of network
+\paragraph{Learning} Learning took place by stepwisely adapting the weights
+$\omega\Ix{read}$ so that the dynamics $t\mapsto r_i(t)$ of the readout
+rates should eventually match a given periodic input pattern $t\mapsto
+a_i(t)$ as accurately as possible. To update $\omega\Ix{read}$ we chose
+a variant of the FORCE algorithm \citep{FORCE} that rapidly modifies the
+feedback loop to keep the errors $|r_i(t) - a_i(t)|$ small from the
+beginning on.
 
-- We chose rate networks in order to apply FORCE rule
-- borrow notation from reward modulation rule paper
-Basic Network dynamics:
-    - Introduce single neuronal network as collection of matrices with specified
-    dynamics
-    - Briefly describe the learning rule (FORCE)
+While the authors of \citep{FORCE} update the weights $\omega\Ix{read}$ in
+fixed intervals (every second time step), we used an adaptive mechanism to
+determine how often learning should occur. This was initially introduced
+for performance reasons, as the single learning steps proved to be the main
+computational bottleneck, but we also found that the long time network
+dynamics tended to be slightly more accurate for adaptive learning steps.
 
-\paragraph{Genes and Challenges}
-    - Introduce parameterized random networks as generators of these networks
-    - Introduce challenges, and the expected success values when paring
-    generators with challenges.
+\paragraph{Parametric Random Networks -- Genotypes} Since we are not
+interested in single network properties but rather want to find out which
+structural elements of networks are important for learning success, we
+introduced \emph{parametric random networks}, or \emph{network genotypes}.
+Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
+which concrete networks are created when $G$ is expressed in a single
+phenotype. Technically $G$ may be understood as a stochastic network
+generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
+value $\omega$.
+The genotypes' parameters $\theta\in\Theta$ were either integer or floating
+point values and can be classified as affecting either (1) the topology of
+$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
+$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
+occupation probability $p$ (for random Erdös-Renyi networks), the
+neighborhood range $k$ (for ring topologies), or the number $l$ of layers
+(for feed-forward networks). We furthermore introduced ratio parameters
+that allowed us to interpolate between different topologies.  Parameters of
+type (2) were the feedback strength \texttt{feedback} (scaling of
+$\omega\Ix{fb}$) and the gain value \texttt{gain} (scaling of
+$\omega\Ix{rec}$ by ).
+
+\paragraph{Fitness} For a given genotype $G$ and a given challenge $C$, the
+fitness $f_C(G)$ of $G$ for $C$ is taken to be the expected value of
+the success rates of $G$'s phenotypes when solving the tasks of the
+challenge $C$. In our simulations, $f_C(G)$ was of course only
+approximated; this was usually done by taking 20 to 100 sample phenotypes
+and sample tasks, where the exact number depended on the standard deviation
+of the previously drawn samples' success rates.
 
@@
 
-\paragraph{Tasks}
+\paragraph{Tasks and Challenges}
 For the optimization to work properly, the problems presented to the networks 
-have (for the most part) to solvable by FORCE learning. While it was shown \cite{}
+have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
 that large recurrent neural networks are capable of learning very complex functions
@@
 
-\begin{thebibliography}{}
-
-\bibitem{mcdermott}
-  Y.-F.~Chen {\it et al.},
-  %``Microwave Photon Counter Based on Josephson Junctions,''
-  Phys.\ Rev.\ Lett.\  {\bf 107}, 217401 (2011).
-
-\end{thebibliography}
+\bibliography{sources}
 

### 2015-08-01 paper/paper.tex
text + script

--- 2015-08-01 paper/paper.tex
@@
 \end{align}
-Algorithmically, the signals are split into chunks which are correlated separately. This saves computational power and 
-prevents minute differences in frequency to accumulate into a huge phase shift, i.e. the phase only has to remain relatively
-constant over the timescale of a single chunk.
+Algorithmically, the signals are split into overlapping chunks which are correlated separately. 
+This saves computational power and prevents minute differences in frequency to accumulate into 
+a huge phase shift, i.e. the phase only has to remain relatively
+constant over the timescale of a single chunk (i.e. about 10 seconds of simulated time).
+
+This measure becomes problematic for functions that are very close to zero for a longer time 
+(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
+different frequencies, this drawback is unproblematic.
 

### 2015-08-01 research/er_gain_feedback.jl
fixed parameter

@@ -6,7 +6,7 @@ ev = EvoNet
 
 
 # here we need to set up our final task set
-ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.5e-2, 5e1))
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.6e-2, 4e0))
 # blacklist all but gain and feedback
 env  = ev.Environment(challenge=ch, blacklist=["size", "percentage"])
 # create the optimizer

### 2015-08-01 src/challenge.jl
logfreqs included

@@ -53,8 +53,9 @@ end
 
 # Helper type
 
-function simple_periodic( f::Function; amplitude=(0.5, 1.5), frequency=(0.025, 2.), args... )
-    return complex_periodic( f, n=1, frequency=frequency, amplitude=amplitude; args... )
+
+function simple_periodic( f::Function; amplitude=(0.5, 1.5), args... )
+    return complex_periodic( f, n=1, amplitude=amplitude; args... )
 end
 
 simple_wave(; args...)  = simple_periodic(x -> sin(2π*x); args...)
@@ -68,6 +69,7 @@ complex_sawtooth(; args...) = complex_periodic(sawtooth_function; args...)
 function complex_periodic( f::Function;
                            n=4, amplitude=(0.75, 1.5), frequency=(0.1, 1.0),
                            phase=(0., 1.),             offset=(-0.2, 0.2),
+                           logfreq=true,
                            clock=false,                clock_frequency=1., # relative freq
                            clock_width=(0.05, 0.1),    clock_amplitude=(0.4, 0.6),
                            clock_phase=(0,1),          clock_offset=(0., 0.) )
@@ -81,7 +83,11 @@ function complex_periodic( f::Function;
     # the wave amplitudes
     amps = Function[ rng -> normal(rng, amplitude...) * 3/(3i) for i in 1:n ]
     # the frequencies
-    freqs = Function[ rng -> unif(rng, frequency...) ]
+    if logfreq
+      freqs = Function[ rng -> exp(unif(rng, map(log, frequency...))) ]
+    else
+      freqs = Function[ rng -> unif(rng, frequency...) ]
+    end
     for i in 2:n push!( freqs, rng -> rand(rng, 0.0:0.25:1) + rand(rng, 1:min(8,2i)) ) end
 
     # the phases

### 2015-08-01 src/challenge.jl
logfreqs included

@@ -84,7 +84,7 @@ function complex_periodic( f::Function;
     amps = Function[ rng -> normal(rng, amplitude...) * 3/(3i) for i in 1:n ]
     # the frequencies
     if logfreq
-      freqs = Function[ rng -> exp(unif(rng, map(log, frequency...))) ]
+      freqs = Function[ rng -> exp(unif(rng, map(log, frequency)...)) ]
     else
       freqs = Function[ rng -> unif(rng, frequency...) ]
     end

### 2015-08-01 paper/paper.tex
tasks

@@
 
+Since we expect/hope that the optimized network extends this range of possible frequencies,
+we choose the frequencies for the optimiztion tasks from the slighly larger interval 
+$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
+for both high and low frequencies.
+
 \paragraph{Assessing the quality of reconstructed functions}

### 2015-08-01 paper/paper.tex
add paper

--- 2015-08-01 paper/paper.tex
@@
     \newcommand{\Ix}[1]{^\mathrm{#1}}
+
+    \newcommand{\param}[1]{\texttt{#1}}
+    \newcommand{\feedback}{\param{feedback}}
+    \newcommand{\gain}{\param{gain}}
+    \newcommand{\size}{N}
+    \newcommand{\ERprob}{p}
+    \newcommand{\RINGk}{k}
+    \newcommand{\FFlayers}{l}
 
@@
 
-\paragraph{Parametric Random Networks -- Genotypes} Since we are not
-interested in single network properties but rather want to find out which
-structural elements of networks are important for learning success, we
-introduced \emph{parametric random networks}, or \emph{network genotypes}.
-Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
-which concrete networks are created when $G$ is expressed in a single
-phenotype. Technically $G$ may be understood as a stochastic network
-generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
-value $\omega$.
-The genotypes' parameters $\theta\in\Theta$ were either integer or floating
-point values and can be classified as affecting either (1) the topology of
-$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
-$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
-occupation probability $p$ (for random Erdös-Renyi networks), the
-neighborhood range $k$ (for ring topologies), or the number $l$ of layers
-(for feed-forward networks). We furthermore introduced ratio parameters
-that allowed us to interpolate between different topologies.  Parameters of
-type (2) were the feedback strength \texttt{feedback} (scaling of
-$\omega\Ix{fb}$) and the gain value \texttt{gain} (scaling of
-$\omega\Ix{rec}$ by ).
-
-\paragraph{Fitness} For a given genotype $G$ and a given challenge $C$, the
-fitness $f_C(G)$ of $G$ for $C$ is taken to be the expected value of
-the success rates of $G$'s phenotypes when solving the tasks of the
-challenge $C$. In our simulations, $f_C(G)$ was of course only
-approximated; this was usually done by taking 20 to 100 sample phenotypes
-and sample tasks, where the exact number depended on the standard deviation
-of the previously drawn samples' success rates.
-
-\paragraph{Genetic Optimization}
-Genetic Optimization:
-
-- Hint / vague description of the codebase
-
-\paragraph{Tasks and Challenges}
+\paragraph{Tasks}
 For the optimization to work properly, the problems presented to the networks 
@@
 using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
-determine the range of $\omega$ in which an (unoptimized) network (i.e. GAIN, 
-FEEDBACK, p, Erdös Renyi) is able to learn the function (see fig. \ref{}).
+determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
+\feedback, \ERprob) is able to learn the function (see fig. \ref{}).
 
-\paragraph{Assessing the quality of reconstructed functions}
+\paragraph{Success}
 In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
@@
 different frequencies, this drawback is unproblematic.
+\paragraph{Parametric Random Networks -- Genotypes} Since we are not
+interested in single network properties but rather want to find out which
+structural elements of networks are important for learning success, we
+introduced \emph{parametric random networks}, or \emph{network genotypes}.
+Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
+which concrete networks are created when $G$ is expressed in a single
+phenotype. Technically $G$ may be understood as a stochastic network
+generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
+value $\omega$.
+The genotypes' parameters $\theta\in\Theta$ were either integer or floating
+point values and can be classified as affecting either (1) the topology of
+$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
+$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
+occupation probability $p$ (for random Erdös-Renyi networks), the
+neighborhood range $k$ (for ring topologies), or the number $l$ of layers
+(for feed-forward networks). We furthermore introduced ratio parameters
+that allowed us to interpolate between different topologies.  Parameters of
+type (2) were the feedback strength $\texttt{feedback}$ (scaling of
+$\omega\Ix{fb}$) and the gain value \gain (scaling of
+$\omega\Ix{rec}$ by ).
+
+\paragraph{Fitness} For a given genotype $G$ and a given challenge $C$, the
+fitness $f_C(G)$ of $G$ for $C$ is taken to be the expected value of
+the success rates of $G$'s phenotypes when solving the tasks of the
+challenge $C$. In our simulations, $f_C(G)$ was of course only
+approximated; this was usually done by taking 20 to 100 sample phenotypes
+and sample tasks, where the exact number depended on the standard deviation
+of the previously drawn samples' success rates.
+
+\paragraph{Genetic Optimization}
+Genetic Optimization:
+
+- Hint / vague description of the codebase
 

### 2015-08-01 paper/paper.tex

@@
 
+\paragraph{Tasks and Challenges}
+For the optimization to work properly, the problems presented to the networks 
+have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
+that large recurrent neural networks are capable of learning very complex functions
+using the FORCE algorithm, the optimization requires a large number of learnings.
+Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
+and as such are less powerful.
+
+To get a simple criterion to estimate whether it might be possible to learn a given function 
+using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
+determine the range of $\omega$ in which an (unoptimized) network (i.e. GAIN, 
+FEEDBACK, p, Erdös Renyi) is able to learn the function (see fig. \ref{}).
+
+Since we expect/hope that the optimized network extends this range of possible frequencies,
+we choose the frequencies for the optimiztion tasks from the slighly larger interval 
+$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
+for both high and low frequencies.
+
+\paragraph{Assessing the quality of reconstructed functions}
+In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
+the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
+$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
+produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.
+
+A measure that is timeshift independent and maximized for functions of identical shape is the 
+maximum of the cross correlation 
+\begin{align}
+f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
+\end{align}
+Algorithmically, the signals are split into overlapping chunks which are correlated separately. 
+This saves computational power and prevents minute differences in frequency to accumulate into 
+a huge phase shift, i.e. the phase only has to remain relatively
+constant over the timescale of a single chunk (i.e. about 10 seconds of simulated time).
+
+This measure becomes problematic for functions that are very close to zero for a longer time 
+(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
+different frequencies, this drawback is unproblematic.
+
 \section{Results}

### 2015-08-01 paper/paper.tex
paper updated

--- 2015-08-01 paper/paper.tex
@@
 
-\paragraph{Tasks}
-For the optimization to work properly, the problems presented to the networks 
-have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
-that large recurrent neural networks are capable of learning very complex functions
-using the FORCE algorithm, the optimization requires a large number of learnings.
-Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
-and as such are less powerful.
+\paragraph{Tasks and Challenges}
+For the learning to work properly, the problems presented to the
+networks have (for the most part) to be solvable by FORCE. While
+it was shown \citep{FORCE} that large recurrent neural networks are capable of
+learning very complex functions, the optimization requires a large number
+of learnings. Therefore, the networks have to be restricted to being quite
+small ($\approx 100$ neurons) and are consequently less powerful.
 
@@
 
+Since we expect/hope that the optimized network extends this range of possible frequencies,
+we choose the frequencies for the optimiztion tasks from the slighly larger interval 
+$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
+for both high and low frequencies.
+
 \paragraph{Success}
-In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
-the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
-$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
-produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.
+In order to optimize the function reconstruction of a force network,
+a measure $f(r', r)$ for the concordance of the target function
+$r'(t)$ and the network readout $r(t)$ is necessary. Simply using the
+difference $f\ix{d}(r, r') = \left\|r-r'\right\|_2$ has the
+drawback that small phase differences could produce huge differences in
+quality, even though $r'$ has the exact same shape as $r$.
 
@@
 \begin{align}
-f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
+    f\ix{c}(r, r') = \max_t \frac{\left<r, r'(\cdot - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
 \end{align}
@@
 a huge phase shift, i.e. the phase only has to remain relatively
-constant over the timescale of a single chunk (i.e. about 10 seconds of simulated time).
+constant over the timescale of a single chunk (about 10 seconds of simulated time).
 

### 2015-08-01 paper/paper.tex
reorder

@@
 Using the above questions as a guideline, we decided to analyze what
-network characteristics makes the FORCE algorithm, introduced in
+network characteristics make the FORCE algorithm, introduced in
 \citep{FORCE}, especially suitable for learning periodic patterns.  More
 precisely, we wanted to identify which properties of a network are
-important for its learning success, and how to choose these properties'
+important for its learning success, and how to choose their
 values for an optimal result.  To this end, one option would be to take an
-initial network and an adequate periodic pattern to be learned, and to
+initial network and an adequate periodic pattern to be learned, and
 optimize the network's internals for this very task. Then patterns may
-emerge that hint at how an optimal network design looks like.
+emerge that hint at how an optimal network design looks.
 
@@
 in the language of genetics, at different \emph{network genotypes}. These
-genotypes contain parametric information about the networks to be created,
-the \emph{phenotypes}. Different phenotypes with the same genotype may thus
+genotypes contain parametric information about the networks (\emph{phenotypes}) to be created. 
+Different phenotypes with the same genotype may thus
 have strong quantitative differences, but they share structural qualities
@@
 problems we faced, we decided to use genetic algorithms for the
-optimization process. Other more direct optimization methods like simulated
-annealing or gradient descent seemed unfeasible for this.
+optimization process.
 
-\section{Concepts and Methods}
-\paragraph{Network Model} We chose the same rate based network architecture
+\section{Concepts and Model}
+\paragraph{Network Model}
+
+MAYBE WE SHOULD MOTIVATE FORCE HERE, WHICH THEN NATURALLY LEADS TO THE CHOICE OF NETWORK BELOW.
+
+We chose the same rate based network architecture
 that is used in the original FORCE publication \citep{FORCE} (denoted as
@@
 from. So we look at a network $\mathcal{N}$ with $N$ internal neurons and
-states $x_i$ for $1 \le i \le N$ that loosely represent the neurons
+states $x_i$ for $1 \le i \le N$ that loosely represent the neurons'
 membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
@@
 
-\paragraph{Tasks and Challenges}
-For the learning to work properly, the problems presented to the
-networks have (for the most part) to be solvable by FORCE. While
-it was shown \citep{FORCE} that large recurrent neural networks are capable of
-learning very complex functions, the optimization requires a large number
-of learnings. Therefore, the networks have to be restricted to being quite
-small ($\approx 100$ neurons) and are consequently less powerful.
-
-To get a simple criterion to estimate whether it might be possible to learn a given function 
-using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
-determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
-\feedback, \ERprob) is able to learn the function (see fig. \ref{}).
-
-Since we expect/hope that the optimized network extends this range of possible frequencies,
-we choose the frequencies for the optimiztion tasks from the slighly larger interval 
-$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
-for both high and low frequencies.
-
-\paragraph{Success}
-In order to optimize the function reconstruction of a force network,
-a measure $f(r', r)$ for the concordance of the target function
-$r'(t)$ and the network readout $r(t)$ is necessary. Simply using the
-difference $f\ix{d}(r, r') = \left\|r-r'\right\|_2$ has the
-drawback that small phase differences could produce huge differences in
-quality, even though $r'$ has the exact same shape as $r$.
-
-A measure that is timeshift independent and maximized for functions of identical shape is the 
-maximum of the cross correlation 
-\begin{align}
-    f\ix{c}(r, r') = \max_t \frac{\left<r, r'(\cdot - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
-\end{align}
-Algorithmically, the signals are split into overlapping chunks which are correlated separately. 
-This saves computational power and prevents minute differences in frequency to accumulate into 
-a huge phase shift, i.e. the phase only has to remain relatively
-constant over the timescale of a single chunk (about 10 seconds of simulated time).
-
-This measure becomes problematic for functions that are very close to zero for a longer time 
-(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
-different frequencies, this drawback is unproblematic.
 \paragraph{Parametric Random Networks -- Genotypes} Since we are not
@@
 
+\section{Methods and Implementation}
+
+\paragraph{Tasks}
+For the learning to work properly, the problems presented to the
+networks have (for the most part) to be solvable by FORCE. While
+it was shown \citep{FORCE} that large recurrent neural networks are capable of
+learning very complex functions, the optimization requires a large number
+of learnings. Therefore, the networks have to be restricted to being quite
+small ($\approx 100$ neurons) and are consequently less powerful.
+
+To get a simple criterion to estimate whether it might be possible to learn a given function 
+using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
+determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
+\feedback, \ERprob) is able to learn the function (see fig. \ref{}).
+
+Since we expect/hope that the optimized network extends this range of possible frequencies,
+we choose the frequencies for the optimization tasks from the slightly larger interval 
+$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
+for both high and low frequencies.
+
+\paragraph{Success}
+In order to optimize the function reconstruction of a force network,
+a measure $f(r', r)$ for the concordance of the target function
+$r'(t)$ and the network readout $r(t)$ is necessary. Simply using the
+difference $f\ix{d}(r, r') = \left\|r-r'\right\|_2$ has the
+drawback that small phase differences could produce huge differences in
+quality, even though $r'$ has the exact same shape as $r$.
+
+A measure that is timeshift independent and maximized for functions of identical shape is the 
+maximum of the cross correlation 
+\begin{align}
+    f\ix{c}(r, r') = \max_t \frac{\left<r, r'(\cdot - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
+\end{align}
+Algorithmically, the signals are split into overlapping chunks which are correlated separately. 
+This saves computational power and prevents minute differences in frequency to accumulate into 
+a huge phase shift, i.e. the phase only has to remain relatively
+constant over the timescale of a single chunk (about 10 seconds of simulated time).
+
+This measure becomes problematic for functions that are very close to zero for a longer time 
+(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
+different frequencies, this drawback is unproblematic.
+
 \paragraph{Genetic Optimization}

### 2015-08-02 paper/paper.tex
motivation for gen alg

@@
 \paragraph{Genetic Optimization}
-Genetic Optimization:
+Due to the stochastic nature of the fitness function $F_C(G)$ of a generator $G$, and the fact that single evaluations
+of $F_C$ are computationally expensive, thus preventing increasing the number of samples such that the estimation $E(F_C)$
+becomes sufficiently smooth, typical optimization methods like gradient descent or simulated annealing are not usable.
+
+ On approach would be to use techniques from Monte Carlo Integration, e.g. stratified or importance sampling, to aid in the
+evaluation of $E(F_C)$, but since $G(\lambda)$ depends noncontinuously on $\lambda$, it can only be applied to the dependence 
+on the challenge $C$. This would mitigate the effects, but not solve the problem.
+
+Therefore, we optimized using a genetic algorithm, which is inherently capable of coping with the stochastic fitness function.
 

### 2015-08-02 paper/paper.tex
message

--- 2015-08-01 paper/paper.tex
@@
 
+\paragraph{Challenges}
+
 \paragraph{Learning} Learning took place by stepwisely adapting the weights
@@
 rates should eventually match a given periodic input pattern $t\mapsto
-a_i(t)$ as accurately as possible. To update $\omega\Ix{read}$ we chose
+r'_i(t)$ as accurately as possible. To update $\omega\Ix{read}$ we chose
 a variant of the FORCE algorithm \citep{FORCE} that rapidly modifies the
-feedback loop to keep the errors $|r_i(t) - a_i(t)|$ small from the
+feedback loop to keep the errors $|r_i(t) - r'_i(t)|$ small from the
 beginning on.

### 2015-08-02 paper/paper.tex
genetic algorithm mk 1

--- 2015-08-02 paper/paper.tex
@@
 
+In each iteration of the algorithm, a generation (set) of $N$ individuals (Generators) $G_t = \left\{ g_i \right\}$ is converted into
+a new one $G_{t+1}$ according to the following scheme:
+\begin{description}
+ \item[Evaluation] The fitness $f_i = F_C(g_i)$ of each individual is mesasured using at least 20 samples (and adaptively more if the $f_i$ are spaced closely, to at msot 100).
+ \item[Selection] All individuals are grouped into disjunct pairs for which we check which one has the higher $f$.
+This is repeated $K=10$ times, counting the wins of each individual.
+Based thereon we use the $p=50\%$ best individuals $S$ to build the next generation.
+
+By using pair comparisons instead of simple ordering according to $f$, we allow weaker individuals to continue into the next generation albeit with low probability), keeping the gene pool more diverse.
+
+\item[Reproduction]
+We generate the next generation $G_{t+1}$ by first generating $2N$ candidates by \emph{mutating} or \emph{recombining} elements of $S$, i.e.
+either changing a single parameter randomly or taking two individuals and randomly choose for each parameter from which parent to take the value.
+We then perform a single sample fitness test and remove all networks that fail, until only $N$ networks are left which form $G_{t+1}$. 
+This way, the most unpromising elements of the new generation are removed before the expensive fitness evaluation. 
+\end{description}
+
 - Hint / vague description of the codebase

### 2015-08-02 paper/paper.tex
use contamination

--- 2015-08-02 paper/paper.tex
@@
 
-A measure that is timeshift independent and maximized for functions of identical shape is the 
+A measure that is time-shift independent and maximized for functions of identical shape is the 
 maximum of the cross correlation 
@@
  On approach would be to use techniques from Monte Carlo Integration, e.g. stratified or importance sampling, to aid in the
-evaluation of $E(F_C)$, but since $G(\lambda)$ depends noncontinuously on $\lambda$, it can only be applied to the dependence 
+evaluation of $E(F_C)$, but since $G(\lambda)$ depends non-continuously on $\lambda$, it can only be applied to the dependence 
 on the challenge $C$. This would mitigate the effects, but not solve the problem.
@@
 \begin{description}
- \item[Evaluation] The fitness $f_i = F_C(g_i)$ of each individual is mesasured using at least 20 samples (and adaptively more if the $f_i$ are spaced closely, to at msot 100).
+ \item[Evaluation] The fitness $f_i = F_C(g_i)$ of each individual is measured using at least 20 samples (and adaptively more if the $f_i$ are spaced closely, to at msot 100).
  \item[Selection] All individuals are grouped into disjunct pairs for which we check which one has the higher $f$.
@@
 
-- Hint / vague description of the codebase
+\paragraph{Choice of Parameters}
 

### 2015-08-02 src/generate/topology.jl
enhanced ring top

@@ -52,6 +52,7 @@ function generate(top::RingTopology, size::Integer, rng::AbstractRNG)
         m[i,j] = rand(rng) < p ? 1 : 0
       else
         m[i,j] = 0
+      end
     end
   end
   return m

### 2015-08-03 paper/paper.tex
planning

--- 2015-08-02 paper/paper.tex
@@
     \newcommand{\Ix}[1]{^\mathrm{#1}}
-
-    \newcommand{\param}[1]{\texttt{#1}}
-    \newcommand{\feedback}{\param{feedback}}
-    \newcommand{\gain}{\param{gain}}
-    \newcommand{\size}{N}
-    \newcommand{\ERprob}{p}
-    \newcommand{\RINGk}{k}
-    \newcommand{\FFlayers}{l}
 
@@
 Using the above questions as a guideline, we decided to analyze what
-network characteristics make the FORCE algorithm, introduced in
+network characteristics makes the FORCE algorithm, introduced in
 \citep{FORCE}, especially suitable for learning periodic patterns.  More
 precisely, we wanted to identify which properties of a network are
-important for its learning success, and how to choose their
+important for its learning success, and how to choose these properties'
 values for an optimal result.  To this end, one option would be to take an
-initial network and an adequate periodic pattern to be learned, and
+initial network and an adequate periodic pattern to be learned, and to
 optimize the network's internals for this very task. Then patterns may
-emerge that hint at how an optimal network design looks.
+emerge that hint at how an optimal network design looks like.
 
@@
 in the language of genetics, at different \emph{network genotypes}. These
-genotypes contain parametric information about the networks (\emph{phenotypes}) to be created. 
-Different phenotypes with the same genotype may thus
+genotypes contain parametric information about the networks to be created,
+the \emph{phenotypes}. Different phenotypes with the same genotype may thus
 have strong quantitative differences, but they share structural qualities
@@
 problems we faced, we decided to use genetic algorithms for the
-optimization process.
+optimization process. Other more direct optimization methods like simulated
+annealing or gradient descent seemed unfeasible for this.
 
-\section{Concepts and Model}
-\paragraph{Network Model}
-
-MAYBE WE SHOULD MOTIVATE FORCE HERE, WHICH THEN NATURALLY LEADS TO THE CHOICE OF NETWORK BELOW.
-
-We chose the same rate based network architecture
+\section{Concepts and Methods}
+\paragraph{Network Model} We chose the same rate based network architecture
 that is used in the original FORCE publication \citep{FORCE} (denoted as
@@
 from. So we look at a network $\mathcal{N}$ with $N$ internal neurons and
-states $x_i$ for $1 \le i \le N$ that loosely represent the neurons'
+states $x_i$ for $1 \le i \le N$ that loosely represent the neurons
 membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
@@
 that allowed us to interpolate between different topologies.  Parameters of
-type (2) were the feedback strength $\texttt{feedback}$ (scaling of
-$\omega\Ix{fb}$) and the gain value \gain (scaling of
+type (2) were the feedback strength \texttt{feedback} (scaling of
+$\omega\Ix{fb}$) and the gain value \texttt{gain} (scaling of
 $\omega\Ix{rec}$ by ).
@@
 
-\section{Methods and Implementation}
+\paragraph{Genetic Optimization}
+Genetic Optimization:
 
-\paragraph{Tasks}
-For the learning to work properly, the problems presented to the
-networks have (for the most part) to be solvable by FORCE. While
-it was shown \citep{FORCE} that large recurrent neural networks are capable of
-learning very complex functions, the optimization requires a large number
-of learnings. Therefore, the networks have to be restricted to being quite
-small ($\approx 100$ neurons) and are consequently less powerful.
+- Hint / vague description of the codebase
+
+\paragraph{Tasks and Challenges}
+For the optimization to work properly, the problems presented to the networks 
+have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
+that large recurrent neural networks are capable of learning very complex functions
+using the FORCE algorithm, the optimization requires a large number of learnings.
+Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
+and as such are less powerful.
 
@@
 using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
-determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
-\feedback, \ERprob) is able to learn the function (see fig. \ref{}).
+determine the range of $\omega$ in which an (unoptimized) network (i.e. GAIN, 
+FEEDBACK, p, Erdös Renyi) is able to learn the function (see fig. \ref{}).
 
 Since we expect/hope that the optimized network extends this range of possible frequencies,
-we choose the frequencies for the optimization tasks from the slightly larger interval 
+we choose the frequencies for the optimiztion tasks from the slighly larger interval 
 $[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
@@
 
-\paragraph{Success}
-In order to optimize the function reconstruction of a force network,
-a measure $f(r', r)$ for the concordance of the target function
-$r'(t)$ and the network readout $r(t)$ is necessary. Simply using the
-difference $f\ix{d}(r, r') = \left\|r-r'\right\|_2$ has the
-drawback that small phase differences could produce huge differences in
-quality, even though $r'$ has the exact same shape as $r$.
+\paragraph{Assessing the quality of reconstructed functions}
+In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
+the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
+$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
+produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.
 
-A measure that is time-shift independent and maximized for functions of identical shape is the 
+A measure that is timeshift independent and maximized for functions of identical shape is the 
 maximum of the cross correlation 
 \begin{align}
-    f\ix{c}(r, r') = \max_t \frac{\left<r, r'(\cdot - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
+f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
 \end{align}
@@
 a huge phase shift, i.e. the phase only has to remain relatively
-constant over the timescale of a single chunk (about 10 seconds of simulated time).
+constant over the timescale of a single chunk (i.e. about 10 seconds of simulated time).
 
@@
 
-\paragraph{Genetic Optimization}
-Due to the stochastic nature of the fitness function $F_C(G)$ of a generator $G$, and the fact that single evaluations
-of $F_C$ are computationally expensive, thus preventing increasing the number of samples such that the estimation $E(F_C)$
-becomes sufficiently smooth, typical optimization methods like gradient descent or simulated annealing are not usable.
+\section{Results}
+Baseline
+gain and feedback tests
 
- On approach would be to use techniques from Monte Carlo Integration, e.g. stratified or importance sampling, to aid in the
-evaluation of $E(F_C)$, but since $G(\lambda)$ depends non-continuously on $\lambda$, it can only be applied to the dependence 
-on the challenge $C$. This would mitigate the effects, but not solve the problem.
+(feedback shifts possible frequencies to higher values)
 
-Therefore, we optimized using a genetic algorithm, which is inherently capable of coping with the stochastic fitness function.
-
-In each iteration of the algorithm, a generation (set) of $N$ individuals (Generators) $G_t = \left\{ g_i \right\}$ is converted into
-a new one $G_{t+1}$ according to the following scheme:
-\begin{description}
- \item[Evaluation] The fitness $f_i = F_C(g_i)$ of each individual is measured using at least 20 samples (and adaptively more if the $f_i$ are spaced closely, to at msot 100).
- \item[Selection] All individuals are grouped into disjunct pairs for which we check which one has the higher $f$.
-This is repeated $K=10$ times, counting the wins of each individual.
-Based thereon we use the $p=50\%$ best individuals $S$ to build the next generation.
-
-By using pair comparisons instead of simple ordering according to $f$, we allow weaker individuals to continue into the next generation albeit with low probability), keeping the gene pool more diverse.
-
-\item[Reproduction]
-We generate the next generation $G_{t+1}$ by first generating $2N$ candidates by \emph{mutating} or \emph{recombining} elements of $S$, i.e.
-either changing a single parameter randomly or taking two individuals and randomly choose for each parameter from which parent to take the value.
-We then perform a single sample fitness test and remove all networks that fail, until only $N$ networks are left which form $G_{t+1}$. 
-This way, the most unpromising elements of the new generation are removed before the expensive fitness evaluation. 
-\end{description}
-
-\paragraph{Choice of Parameters}
-
-\section{Results}
+success of omega: erdös-renyi, ring
+optimizations: ER topology, Ring Topology, Mixing
 

### 2015-08-03 paper/paper.tex
Paper chapters 1,2 ready, 3 nearly

--- 2015-08-02 paper/paper.tex
@@
 
-To address these issues we looked at \emph{parameterized random networks}, or
-in the language of genetics, at different \emph{network genotypes}. These
-genotypes contain parametric information about the networks (\emph{phenotypes}) to be created. 
-Different phenotypes with the same genotype may thus
-have strong quantitative differences, but they share structural qualities
-whose fitness for a preassigned task we wanted to analyze and improve.
+To address these issues we looked at parameterized random networks, or in
+the language of genetics, at different \emph{network genotypes}. These
+genotypes contain parametric information about the networks
+(\emph{phenotypes}) to be created. 
+Different phenotypes with the same genotype may thus have strong
+quantitative differences, but they share structural qualities whose fitness
+for a preassigned range of tasks we wanted to analyze and improve.
 
@@
 
+We first lay out the main aspects of our model in a 
+partially abstract fashion in the next section.
+%rather high-level fashion. 
+More details on the actual implementation are then provided in the third
+section (or ultimately in the published source
+code / supplementary material).
+Finally, some of the obtained results are presented and discussed.
+
 \section{Concepts and Model}
+\label{sec:concepts}
 \paragraph{Network Model} 
-
-MAYBE WE SHOULD MOTIVATE FORCE HERE, WHICH THEN NATURALLY LEADS TO THE CHOICE OF NETWORK BELOW.
+%MAYBE WE SHOULD MOTIVATE FORCE HERE, WHICH THEN NATURALLY LEADS TO THE CHOICE OF NETWORK BELOW.
 
@@
 membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
-given by $r_i = \tanh x_i$.  Further, the network contains $R$ readout
+given by $r_i = \tanh x_i$.  Furthermore, the network contains $R$ readout
 neurons with states $z_i$ for $1 \le i \le R$, where
 \begin{equation} 
-    z_i = \sum_{j=1}^{N}\omega\Ix{read}_{ij}\,r_j~.
+    z_i = \sum_{j=1}^{N}\omega\Ix{read}_{ij}\,r_j
 \end{equation}
+with the readout weights $\omega\Ix{read}$.
 Taking into account both internal dynamics as well as an external feedback
@@
 a simple Newton method with integration time steps $dt = 0.01$, which was
-also used in \citep{FORCE}. (How we constructed $\omega\Ix{rec}$ and
-$\omega\Ix{fb}$ will be discussed in \emph{Parametric Random Networks})
+also used in \citep{FORCE}. 
+%(How we constructed $\omega\Ix{rec}$ and
+%$\omega\Ix{fb}$ will be described in \emph{Parametric Random Networks})
 
@@
 $\omega\Ix{read}$ so that the dynamics $t\mapsto r_i(t)$ of the readout
-rates should eventually match a given periodic input pattern $t\mapsto
-a_i(t)$ as accurately as possible. To update $\omega\Ix{read}$ we chose
-a variant of the FORCE algorithm \citep{FORCE} that rapidly modifies the
-feedback loop to keep the errors $|r_i(t) - a_i(t)|$ small from the
-beginning on.
+neurons should eventually match a given periodic target pattern $t\mapsto
+r'_i(t)$, called a \emph{task} from now on, as accurately as possible. To
+update $\omega\Ix{read}$ we chose a variant of the FORCE algorithm
+\citep{FORCE} that rapidly modifies the feedback loop to keep the errors
+$|r_i(t) - r'_i(t)|$ small from the beginning on.
 
-While the authors of \citep{FORCE} update the weights $\omega\Ix{read}$ in
+While the authors of \citep{FORCE} update the weights $\lambda\Ix{read}$ in
 fixed intervals (every second time step), we used an adaptive mechanism to
@@
 
-\paragraph{Parametric Random Networks -- Genotypes} Since we are not
+\paragraph{Network Genotypes} Since we are not
 interested in single network properties but rather want to find out which
 structural elements of networks are important for learning success, we
-introduced \emph{parametric random networks}, or \emph{network genotypes}.
+introduced  \emph{network genotypes}.
 Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
@@
 phenotype. Technically $G$ may be understood as a stochastic network
-generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
-value $\omega$.
+generator that returns a network $\mathcal{N}=G(\lambda)$ for every seed
+value $\lambda$.
 The genotypes' parameters $\theta\in\Theta$ were either integer or floating
@@
 (for feed-forward networks). We furthermore introduced ratio parameters
-that allowed us to interpolate between different topologies.  Parameters of
-type (2) were the feedback strength $\texttt{feedback}$ (scaling of
-$\omega\Ix{fb}$) and the gain value \gain (scaling of
-$\omega\Ix{rec}$ by ).
+that allowed us to interpolate between different topologies. Optimized
+parameters of type (2) were the feedback strength $\feedback$ (with
+$\omega\Ix{fb}\propto \feedback$) and the gain value $\gain$ (with
+$\omega\Ix{rec} \propto \gain$).
 
-\paragraph{Fitness} For a given genotype $G$ and a given challenge $C$, the
-fitness $f_C(G)$ of $G$ for $C$ is taken to be the expected value of
-the success rates of $G$'s phenotypes when solving the tasks of the
-challenge $C$. In our simulations, $f_C(G)$ was of course only
-approximated; this was usually done by taking 20 to 100 sample phenotypes
-and sample tasks, where the exact number depended on the standard deviation
-of the previously drawn samples' success rates.
+\paragraph{Fitness} The next step was to define the \emph{fitness} of a given
+genotype when confronted with a diverse range of tasks, like waves of
+different frequencies. To express this formally, we introduced
+\emph{challenges}: For a given seed value $\lambda$, the challenge $C$
+yields a task $r' = C(\lambda)$ according to some (parametrized)
+distribution. If we now make use of a success function $S(\mathcal{N},
+r')\in\{0, 1\}$ that decides, if the network $\mathcal{N}$
+has successfully reproduced the desired output $r'$, the fitness $F_C(G)$
+of a genotype $G$ for a challenge $C$ may reasonably be defined as the
+expectation value of $\lambda \mapsto S(G(\lambda), C(\lambda))$. For fixed
+challenges, the main task of optimizing the genotype is now reduced to
+finding parameters $\Theta$ that maximize $F_C$.
 
+% IDEA: Maybe put this part at the end of the paper. Most things needed to
+% understand the results (in principle) have already been mentioned at this
+% point.
 \section{Methods and Implementation}
+\label{sec:implementation}
 
 \paragraph{Tasks}
-For the learning to work properly, the problems presented to the
-networks have (for the most part) to be solvable by FORCE. While
-it was shown \citep{FORCE} that large recurrent neural networks are capable of
-learning very complex functions, the optimization requires a large number
-of learnings. Therefore, the networks have to be restricted to being quite
-small ($\approx 100$ neurons) and are consequently less powerful.
+For the optimization to work properly, the tasks presented to the networks
+must be reasonably chosen; they must be learnable by FORCE (at least in
+principle) and must be compatible with technical aspects like the
+integration time step.
+%have (for the most part) to be solvable by FORCE. 
+While it was shown in \citep{FORCE} that large recurrent neural networks are
+capable of learning very complex tasks, the optimization requires an
+extensive number of learning processes, since every single evaluation of
+the fitness $F_C$ needs dozens of tested phenotypes (see below). Therefore,
+the networks had to be restricted to being quite small ($\approx 100$
+neurons) and are consequently less powerful.
 
-To get a simple criterion to estimate whether it might be possible to learn a given function 
-using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
-determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
-\feedback, \ERprob) is able to learn the function (see fig. \ref{}).
+To get a first estimate whether it might be possible to learn a given task
+using FORCE with 100 neurons in our network model, we looked at a single
+sinusoidal of frequency $\omega$ and determined the frequency range in
+which an (unoptimized) network (i.e. \gain, \feedback, \ERprob) is able to
+learn the function (see fig. \ref{}).
 
-Since we expect/hope that the optimized network extends this range of possible frequencies,
-we choose the frequencies for the optimization tasks from the slightly larger interval 
-$[0.6e-2, 4]$ such that their logarithms are uniformly spaced. This allows for good resolution 
-for both high and low frequencies.
+Since we expect/hope that the optimized network extends this range of
+possible frequencies, we choose the frequencies for the optimization tasks
+from the slightly larger interval $[0.6e-2, 4]$ such that their logarithms
+are uniformly spaced. This allows for good resolution for both high and low
+frequencies.
 
 \paragraph{Success}
-In order to optimize the function reconstruction of a force network,
-a measure $f(r', r)$ for the concordance of the target function
-$r'(t)$ and the network readout $r(t)$ is necessary. Simply using the
-difference $f\ix{d}(r, r') = \left\|r-r'\right\|_2$ has the
-drawback that small phase differences could produce huge differences in
-quality, even though $r'$ has the exact same shape as $r$.
+As mentioned in \emph{Concepts and Methods}, we need a function $S$
+that decides weather a network succeeded or failed to solve a task.
+To this end
+%In order to optimize the function reconstruction of a force network,
+a measure $f(r', r)$ for the concordance of the target function $r'(t)$ and
+the network readout $r(t)$ is sought. Though one could use a simple
+difference based metric, like the norm $f\ix{2}(r, r')
+= \left\|r-r'\right\|_2$, this has the drawback that small phase
+mismatches could produce high differences, even though the
+shape of $r$ matches the one of $r'$ well.
 
-A measure that is time-shift independent and maximized for functions of identical shape is the 
-maximum of the cross correlation 
+Instead, a measure that is time-shift independent and maximized for functions of
+identical shape is the maximum of the cross correlation 
 \begin{align}
-    f\ix{c}(r, r') = \max_t \frac{\left<r, r'(\cdot - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
+    f\ix{cor}(r, r') = \max_t \frac{\left<r, r'(\cdot
+    - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
 \end{align}
-Algorithmically, the signals are split into overlapping chunks which are correlated separately. 
-This saves computational power and prevents minute differences in frequency to accumulate into 
-a huge phase shift, i.e. the phase only has to remain relatively
-constant over the timescale of a single chunk (about 10 seconds of simulated time).
+Algorithmically, the signals $r$ and $r'$ are split into overlapping chunks
+which are correlated separately.  This saves computational power and
+prevents minute differences in frequency to accumulate into a huge phase
+shift, so the phase only has to remain relatively constant over the
+timescale of a single chunk (which is about 10 seconds of simulated time).
+This measure becomes problematic for functions that are very close to zero
+for a longer time (order of chunk size), but since we are going to work
+with superpositions of sinusoidals of different frequencies, this drawback
+is unproblematic.
 
-This measure becomes problematic for functions that are very close to zero for a longer time 
-(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
-different frequencies, this drawback is unproblematic.
+Using this measure of quality, the success function for a network
+$\mathcal{N}$ with readout $r$ for the task $r'$ was taken to be 
+\begin{equation}
+    S(\mathcal{N}, r') = \begin{cases} 1\quad\mathrm{if}~f\ix{cor}(r, r') \ge 0.95,\\
+                                       0\quad\mathrm{if}~f\ix{cor}(r, r') < 0.95~.
+                         \end{cases}
+\end{equation}
+The threshold value of 0.95 was determined empirically, as signals $r$ with
+$f(r, r') \ge 0.95$ met our assessment of what we considered a good
+reproduction of the target pattern.
 
 \paragraph{Genetic Optimization}
-Due to the stochastic nature of the fitness function $F_C(G)$ of a generator $G$, and the fact that single evaluations
-of $F_C$ are computationally expensive, thus preventing increasing the number of samples such that the estimation $E(F_C)$
-becomes sufficiently smooth, typical optimization methods like gradient descent or simulated annealing are not usable.
+Due to the stochastic nature of the fitness function $F_C(G)$ of a genotype
+$G$, a very high number of samples would be required to obtain sufficiently
+smooth approximations of $F_C(G)$ for typical optimization methods like
+gradient descend or simulated annealing. But as the computational cost for
+even one learning process was considerable high, these methods were not
+feasible.
 
- On approach would be to use techniques from Monte Carlo Integration, e.g. stratified or importance sampling, to aid in the
-evaluation of $E(F_C)$, but since $G(\lambda)$ depends non-continuously on $\lambda$, it can only be applied to the dependence 
-on the challenge $C$. This would mitigate the effects, but not solve the problem.
+%and the fact that single evaluations of $F_C$ are
+%computationally very expensive, thus preventing increasing the number of
+%samples such that the estimation $E(F_C)$ becomes sufficiently smooth,
+%typical optimization methods like gradient descent or simulated annealing
+%are not usable.
 
-Therefore, we optimized using a genetic algorithm, which is inherently capable of coping with the stochastic fitness function.
+An alternative approach would be to use techniques from Monte Carlo
+integration, e.g.  stratified or importance sampling, to aid in the
+evaluation of $F_C(G)$, but since $G(\lambda)$ and $C(\lambda)$ depend
+non-continuously on the seed value $\lambda$, this too would not solve the
+problem.
 
-In each iteration of the algorithm, a generation (set) of $N$ individuals (Generators) $G_t = \left\{ g_i \right\}$ is converted into
-a new one $G_{t+1}$ according to the following scheme:
+Therefore, we optimized $F_C(G)$ using a genetic algorithm, which is inherently
+capable of coping with the stochastic fitness function. In each iteration
+of the algorithm, a generation $\mathcal{G}_t = \left\{ G_i~|~1\le i\le N\right\}$
+of $N$ individuals (Generators) is converted into a new one, $\mathcal{G}_{t+1}$,
+according to the following scheme:
 \begin{description}
- \item[Evaluation] The fitness $f_i = F_C(g_i)$ of each individual is measured using at least 20 samples (and adaptively more if the $f_i$ are spaced closely, to at msot 100).
- \item[Selection] All individuals are grouped into disjunct pairs for which we check which one has the higher $f$.
-This is repeated $K=10$ times, counting the wins of each individual.
-Based thereon we use the $p=50\%$ best individuals $S$ to build the next generation.
+ \item[Evaluation] The fitness $F_i = F_C(G_i)$ of each individual is
+     measured using at least 20 samples (and adaptively more if the
+     $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
+ \item[Selection] All individuals are grouped into disjunct pairs for which
+     we check which one has the higher fitness. This is repeated $k=10$ times,
+     counting the number of wins for each individual. Based thereon we use the
+     $p=50\%$ best individuals $\mathcal{G}\Ix{survive}_t$ to build the next
+     generation.  By using pair comparisons instead of simple ordering
+     according to the fitness, we allow weaker individuals to reach
+     the next generation (albeit with low probability), keeping the gene
+     pool more diverse.
 
-By using pair comparisons instead of simple ordering according to $f$, we allow weaker individuals to continue into the next generation albeit with low probability), keeping the gene pool more diverse.
-
-\item[Reproduction]
-We generate the next generation $G_{t+1}$ by first generating $2N$ candidates by \emph{mutating} or \emph{recombining} elements of $S$, i.e.
-either changing a single parameter randomly or taking two individuals and randomly choose for each parameter from which parent to take the value.
-We then perform a single sample fitness test and remove all networks that fail, until only $N$ networks are left which form $G_{t+1}$. 
-This way, the most unpromising elements of the new generation are removed before the expensive fitness evaluation. 
+\item[Reproduction] We obtain the next generation $\mathcal{G}_{t+1}$ by first
+    generating $2N$ candidates via \emph{mutating} or \emph{recombining}
+    elements of $\mathcal{G}\Ix{survive}_t$. This means we either change
+    a single parameter randomly or take two individuals and randomly choose
+    for each parameter from which parent to take the value.  We then
+    perform a single sample fitness test and remove all networks that fail,
+    until only $N$ networks are left; these form $\mathcal{G}_{t+1}$. By
+    this approach the most unpromising elements of the new generation are
+    removed before the expensive fitness evaluation takes place. 
 \end{description}
@@
 \section{Results}
+Baseline
+gain and feedback tests
+
+(feedback shifts possible frequencies to higher values)
+
+success of omega: erdös-renyi, ring
+optimizations: ER topology, Ring Topology, Mixing
 

### 2015-08-04 src/optimize/genetic_optimizer.jl
removed old record population

@@ -126,8 +126,7 @@ function fight_till_death( opt::GeneticOptimizer, population::Vector{AbstractGen
   end
 
   # it is not really nice to record here :(
-  record_population_(opt.recorder, population, success, opt.generation)
-  #=record_population(opt.recorder, population, success, opt.generation)=#
+  record_population(opt.recorder, population, success, opt.generation)
 
   wins = zeros(length(population))
   NUM_FIGHTS = 10
@@ -172,7 +171,7 @@ function mean_success(suc::Vector{AbstractRating})
   return mean::Float64, sqrt(variance)::Float64
 end
 
-function record_population_(rec::Recorder, pop::Vector{AbstractGenerator}, suc::Vector{AbstractRating}, generation::Integer)
+function record_population(rec::Recorder, pop::Vector{AbstractGenerator}, suc::Vector{AbstractRating}, generation::Integer)
   # get all parameters that occur for the generators
   #
   for i = 1:length(pop)
@@ -186,25 +185,6 @@ function record_population_(rec::Recorder, pop::Vector{AbstractGenerator}, suc::
   end
 end
 
-# writes info about a population
-function record_population(rec::Recorder, pop::Vector{AbstractGenerator}, suc::Vector{AbstractRating}, generation::Integer)
-  # collect info about all gens
-  for i = 1:length(pop)
-    succ = Float64[suc[i].quota, suc[i].quality, suc[i].timeshift]
-    pars = Float64[]
-    # UGLY UGLY UGLY
-    for p in export_params(pop[i])
-      if isa(p.val, Real)
-        pars = vcat(pars, p.val)
-      elseif isa(p.val, Vector{Float64})
-        pars = vcat(pars, p.val...)
-      else
-        @assert false "$(typeof(p.val))"
-      end
-    end
-    record(rec, 1, vcat([generation], succ, pars))
-  end
-end
 
 # performs recombination and mutation / currently mutually exclusive
 function calculate_next_generation( opt::GeneticOptimizer, parents::Vector{AbstractGenerator}, N::Integer)

### 2015-08-04 src/optimize/genetic_optimizer.jl
record survival too

@@ -125,9 +125,6 @@ function fight_till_death( opt::GeneticOptimizer, population::Vector{AbstractGen
     end
   end
 
-  # it is not really nice to record here :(
-  record_population(opt.recorder, population, success, opt.generation)
-
   wins = zeros(length(population))
   NUM_FIGHTS = 10
   for i = 1:NUM_FIGHTS
@@ -145,12 +142,17 @@ function fight_till_death( opt::GeneticOptimizer, population::Vector{AbstractGen
   end
 
   num_survivors = ceil(length(population) * (1.0 - reduction_rate) )
+  survived = zeros( length(population) )
   survivors = AbstractGenerator[]
   for i = 1:num_survivors
     index = indmax(wins)
     push!(survivors, population[index]  )
     wins[index] = 0 # prevent double selection
+    survived[index] = 1
   end
+
+  # record the population, includin who survived
+  record_population(opt.recorder, population, success, survived, opt.generation)
 
   return shuffle(rng, survivors) # make sure they are not ordered in any particular way
 end
@@ -171,14 +173,15 @@ function mean_success(suc::Vector{AbstractRating})
   return mean::Float64, sqrt(variance)::Float64
 end
 
-function record_population(rec::Recorder, pop::Vector{AbstractGenerator}, suc::Vector{AbstractRating}, generation::Integer)
+function record_population(rec::Recorder, pop::Vector{AbstractGenerator}, suc::Vector{AbstractRating}, surv::Vector, generation::Integer)
   # get all parameters that occur for the generators
   #
   for i = 1:length(pop)
-    record(rec, "G", generation)
+    record(rec, "G",  generation)
     record(rec, "QT", suc[i].quota)
     record(rec, "QL", suc[i].quality)
     record(rec, "TS", suc[i].timeshift)
+    record(rec, "SV", surv[i])
     for p in export_params(pop[i])
       record(rec, p.name, p.val)
     end
@@ -259,7 +262,7 @@ end
 
 function save_evolution(file, opt::GeneticOptimizer)
   names = Param.get_parameter_names(opt.population[1])
-  output = [opt.recorder["G"] opt.recorder["QT"] opt.recorder["QL"] opt.recorder["TS"]]
[…]

### 2015-08-04 research/ring_vs_er.jl
fix

@@ -19,14 +19,14 @@ k = 2
 ertop = ev.ErdösRenyiTopology( p )
 ringtop = ev.RingTopology( k )
 # choose gain and feedback as in the paper
-gen  = ev.SparseFRGenerator( N, topology = ev.MetaTopology(ertop, ringtop), gain = 1.2, feedback = 2 )
+gen  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop]), gain = 1.2, feedback = 2 )
 ev.init_population!(gopt, [(gen, 0.5)], 50)
 
 if N != 100
   info("this seems to be a test run. use N=100 to create final results")
 end
 
-for i = 1:25
+for i = 1:50
   ev.step!(gopt)
   ev.save_evolution("ring_vs_er.dat", gopt)
 end

### 2015-08-04 src/optimize/genetic_optimizer.jl
increased minimum number of samples

@@ -98,11 +98,11 @@ end
 # lets the members of a population fight
 function fight_till_death( opt::GeneticOptimizer, population::Vector{AbstractGenerator};
                           rng::AbstractRNG = opt.rng, compare::Function = opt.compare,
-                          reduction_rate::Float64 = 0.5, max_samples = 100
+                          reduction_rate::Float64 = 0.5, max_samples::Integer = 100, 
+                          samples::Integer = 40
                           )
    # collection of all generators that survive
-  success = rate_population_parallel(opt, seed=randseed(rng), pop = population, samples = 20)
-  samples = 20
+  success = rate_population_parallel(opt, seed=randseed(rng), pop = population, samples = samples)
   mean, stddev = mean_success(success)
   while(samples < max_samples)
     req1 = ceil(4*(mean * (1-mean) / stddev)^2) # first criterion: error

### 2015-08-04 research/high_frq.jl
high frquency test

@@ -1,0 +1,32 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+ev = EvoNet
+#optimize secondary parameters (feedback, gain, percentage)
+
+
+# here we need to set up our final task set
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.3, 3e0))
+# blacklist all but gain and feedback
+env  = ev.Environment(challenge=ch, blacklist=["size"])
+# create the optimizer
+gopt = ev.GeneticOptimizer( ev.fitness_in_environment, ev.compare_fitness, env=env )
+
+# initialize the population by using the
+N = 100 # use 100 for final run
+p = 0.1
+ertop = ev.ErdösRenyiTopology(p)
+# choose gain and feedback far from the values of the original paper, to show 
+# that the optimization works.
+gen  = ev.SparseFRGenerator( N, topology = ertop, gain = 1.2, feedback = 2 )
+ev.init_population!(gopt, gen, 50)
+
+if N != 100
+  info("this seems to be a test run. use N=100 to create final results")
+end
+
+for i = 1:50
+  ev.step!(gopt)
+  ev.save_evolution("high_frq_opt.dat", gopt)
+end
+

### 2015-08-04 research/low_frq.jl
low frq test

@@ -1,0 +1,32 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+ev = EvoNet
+#optimize secondary parameters (feedback, gain, percentage)
+
+
+# here we need to set up our final task set
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.02, 0.02))
+# blacklist all but gain and feedback
+env  = ev.Environment(challenge=ch, blacklist=["size"])
+# create the optimizer
+gopt = ev.GeneticOptimizer( ev.fitness_in_environment, ev.compare_fitness, env=env )
+
+# initialize the population by using the
+N = 100 # use 100 for final run
+p = 0.1
+ertop = ev.ErdösRenyiTopology(p)
+# choose gain and feedback far from the values of the original paper, to show 
+# that the optimization works.
+gen  = ev.SparseFRGenerator( N, topology = ertop, gain = 1.2, feedback = 2 )
+ev.init_population!(gopt, gen, 50)
+
+if N != 100
+  info("this seems to be a test run. use N=100 to create final results")
+end
+
+for i = 1:50
+  ev.step!(gopt)
+  ev.save_evolution("low_frq_opt.dat", gopt)
+end
+

### 2015-08-05 research/full_frq.jl
full freq opt

@@ -1,0 +1,32 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+ev = EvoNet
+#optimize secondary parameters (feedback, gain, percentage)
+
+
+# here we need to set up our final task set
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.02, 3e0))
+# blacklist all but gain and feedback
+env  = ev.Environment(challenge=ch, blacklist=["size"])
+# create the optimizer
+gopt = ev.GeneticOptimizer( ev.fitness_in_environment, ev.compare_fitness, env=env )
+
+# initialize the population by using the
+N = 100 # use 100 for final run
+p = 0.1
+ertop = ev.ErdösRenyiTopology(p)
+# choose gain and feedback far from the values of the original paper, to show 
+# that the optimization works.
+gen  = ev.SparseFRGenerator( N, topology = ertop, gain = 1.2, feedback = 2 )
+ev.init_population!(gopt, gen, 50)
+
+if N != 100
+  info("this seems to be a test run. use N=100 to create final results")
+end
+
+for i = 1:100
+  ev.step!(gopt)
+  ev.save_evolution("full_frq_opt.dat", gopt)
+end
+

### 2015-08-05 research/low_frq.jl
fix frequency range

@@ -6,7 +6,7 @@ ev = EvoNet
 
 
 # here we need to set up our final task set
-ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.02, 0.02))
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.02, 0.002))
 # blacklist all but gain and feedback
 env  = ev.Environment(challenge=ch, blacklist=["size"])
 # create the optimizer
@@ -25,7 +25,7 @@ if N != 100
   info("this seems to be a test run. use N=100 to create final results")
 end
 
-for i = 1:50
+for i = 1:100
   ev.step!(gopt)
   ev.save_evolution("low_frq_opt.dat", gopt)
 end

### 2015-08-05 research/low_frq.jl
not so unfair

@@ -6,7 +6,7 @@ ev = EvoNet
 
 
 # here we need to set up our final task set
-ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.02, 0.002))
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.04, 0.004))
 # blacklist all but gain and feedback
 env  = ev.Environment(challenge=ch, blacklist=["size"])
 # create the optimizer

### 2015-08-05 research/q_of_omega_high.jl
added q_of_omega_{unoptimized, high}.jl

@@ -1,0 +1,23 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+#=ev = EvoNet=#
+
+# F_C(ω) curve for an unoptimized genotype
+
+@everywhere function rating(f)
+  N = 100 # use 100 for final run
+  p = .0998625185385132
+  gain = 1.1356512394227996
+  feedback = 10.985208608245522
+  ertop = EvoNet.ErdösRenyiTopology(p)
+  gen  = EvoNet.SparseFRGenerator( N, topology = ertop, gain = gain, feedback = feedback )
+  ch   = EvoNet.simple_wave(amplitude=1, frequency=f, offset=0)
+  env  = EvoNet.Environment(challenge=ch)
+  s = EvoNet.fitness_in_environment(gen, samples=200, env=env)
+  return EvoNet.get_value(s)
+end
+
+freq = logspace(-3, 0.5, 1000)
+result = pmap(x->rating(x), freq)
+writedlm("freqrange_high.txt", [freq result])

### 2015-08-05 research/q_of_omega_unoptimized.jl

@@ -1,0 +1,23 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+#=ev = EvoNet=#
+
+# F_C(ω) curve for an unoptimized genotype
+
+@everywhere function rating(f)
+  N = 100 # use 100 for final run
+  p = 0.1
+  gain = 1.5
+  feedback = 2
+  ertop = EvoNet.ErdösRenyiTopology(p)
+  gen  = EvoNet.SparseFRGenerator( N, topology = ertop, gain = gain, feedback = feedback )
+  ch   = EvoNet.simple_wave(amplitude=1, frequency=f, offset=0)
+  env  = EvoNet.Environment(challenge=ch)
+  s = EvoNet.fitness_in_environment(gen, samples=200, env=env)
+  return EvoNet.get_value(s)
+end
+
+freq = logspace(-3, 0.5, 200)
+result = pmap(x->rating(x), freq)
+writedlm("freqrange_unoptimized.txt", [freq result])

### 2015-08-05 research/full_frq.jl
data

@@ -25,7 +25,7 @@ if N != 100
   info("this seems to be a test run. use N=100 to create final results")
 end
 
-for i = 1:100
+for i = 1:200
   ev.step!(gopt)
   ev.save_evolution("full_frq_opt.dat", gopt)
 end

### 2015-08-05 paper/paper.tex
paper updated
first picture included

@@
     \usepackage{blindtext}
+    \usepackage{enumitem}
 
@@
 	\usepackage[colorlinks,hyperindex]{hyperref}
+    \usepackage{color}
 	\hypersetup
@@
 We first lay out the main aspects of our model in a 
-partially abstract fashion in the next section.
-%rather high-level fashion. 
+%partially abstract fashion in the next section.
+rather high-level overview. 
 More details on the actual implementation are then provided in the third
@@
 according to the following scheme:
-\begin{description}
+\begin{description}[itemsep=1ex, leftmargin=0pt]
  \item[Evaluation] The fitness $F_i = F_C(G_i)$ of each individual is
@@
 \paragraph{Choice of Parameters}
+- $N=100$ fixed, since greater $N$ always increased quality and computational cost
+increased quadraticaly to $N$
+- Mutation rate
 
 \section{Results}
+\subsection{Frequency Dependent Optimizations}
+- short description: get feeling for frequency range (by using unoptimized parameters
+from the force paper), make optimizations with 3 different challenges: 1) challenges that cover the frequency area completely, challenges that cover low frequencies, and challanges that cover high frequencies, then pick genotypes with high fitness.
+\subsection{Topology Optimizations}
 Baseline
@@
 
+\begin{figure}
+    {\footnotesize\input{pictures/freqrange}}
+    \caption{CAPTION TEXT}
+    \label{fig:freqrange}
+\end{figure}
+
 \section{Discussion}
+- Remarks
+    * Influence of initial conditions
+        -> Degenerated initial population poses major problems (does not
+        find same / takes very long to find same parameters as when started
+        with diversity).
+        -> With broad diversity: often fast convergence -> Gen. Optimizer
+        good as optimizing filter, but has hard times to explore new
+        parameter regions if no right genotypes are present (especially isolated
+        maxima of $F_C$). Solution: Could introduce migrants, that supply
+        the gene-pool with underrepresented genotypes.
+    * According to experience: maybe not ideal for finding single set of
+      parameters but rather for successful parameter-ranges (branches?)
+
+- Problems
+    * Many magic parameters scattered in the model that were not examined exhaustively
+    * Only few and uncomprehensive number of test / experiments conducted due to the
+      computational cost.
+    * Test with several other topologies did not yield concrete results,
+      only rough tendencies.
+    * Influence of statistics: Learning success depends stronger on choice
+      of challenges than parameters (partially) -> must be very careful how to
+      interpret the results
 

### 2015-08-05 research/q_of_omega_low.jl
q_of_omega_low added

@@ -7,9 +7,9 @@ using EvoNet
 
 @everywhere function rating(f)
   N = 100 # use 100 for final run
-  p = .11807608106235096
-  gain = 1.288355037230867	
-  feedback = .7798470419161503
+  p = 
+  gain = 
+  feedback = 
   ertop = EvoNet.ErdösRenyiTopology(p)
   gen  = EvoNet.SparseFRGenerator( N, topology = ertop, gain = gain, feedback = feedback )
   ch   = EvoNet.simple_wave(amplitude=1, frequency=f, offset=0)
@@ -18,6 +18,6 @@ using EvoNet
   return EvoNet.get_value(s)
 end
 
-freq = logspace(-3, 0.5, 1000)
+freq = logspace(-3, 0.5, 200)
 result = pmap(x->rating(x), freq)
 writedlm("freqrange_unoptimized.txt", [freq result])

### 2015-08-05 research/q_of_omega_low.jl
updated

@@ -7,9 +7,9 @@ using EvoNet
 
 @everywhere function rating(f)
   N = 100 # use 100 for final run
-  p = 
-  gain = 
-  feedback = 
+  p = .11807608106235096
+  gain = 1.288355037230867	
+  feedback = .7798470419161503
   ertop = EvoNet.ErdösRenyiTopology(p)
   gen  = EvoNet.SparseFRGenerator( N, topology = ertop, gain = gain, feedback = feedback )
   ch   = EvoNet.simple_wave(amplitude=1, frequency=f, offset=0)

### 2015-08-05 research/q_of_omega_low.jl
updated

@@ -18,6 +18,6 @@ using EvoNet
   return EvoNet.get_value(s)
 end
 
-freq = logspace(-3, 0.5, 200)
+freq = logspace(-3, 0.5, 1000)
 result = pmap(x->rating(x), freq)
 writedlm("freqrange_unoptimized.txt", [freq result])

### 2015-08-05 src/param/meta_parametric_object.jl
new meta parameter change

@@ -12,15 +12,29 @@ end
 
 # Meta Generator relative generator contribution
 function random_param( param::MetaCombinationParam, rng::AbstractRNG; s::Real = 0.1 )
-  new_random_val = Float64[rand(rng) for i in param.val]
-  new_random_val /= sum(new_random_val)
+  # choose two paramerters
+  p1 = rand(rng, 1:length(param.val))
+  p2 = rand(rng, 1:length(param.val))
 
-  # Obtain new parameter values by relative changes of +-0.1*s
-  new_val = param.val + s*new_random_val
-  new_val /= sum(new_val)
+  v1 = param.val[p1]
+  v2 = param.val[p2]
+  sum = v1 + v2
+  # rescale to 1
+  v1 /= sum
+  v2 /= sum
 
-  # Check if the new value is within the boundaries and return new param
-  return MetaCombinationParam(param.name, new_val)
+  # select a change rate
+  change = min(v1, 1-v2, abs(randn(rng) * s))
+
+  v1 += change
+  v2 -= change
+
+  # scale back
+  v1 *= sum
+  v2 *= sum
+
+  param.val[p1] = v1
+  param.val[p2] = v2
 end
 
 #####################################################################################

### 2015-08-05 research/er_ring.jl
fixes + er ring

@@ -1,0 +1,34 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+ev = EvoNet
+#optimize secondary parameters (feedback, gain)
+
+
+# here we need to set up our final task set
+ch   = EvoNet.simple_wave(amplitude=1, frequency=(0.6e-2, 4e0))
+# blacklist all but gain and feedback
+env  = ev.Environment(challenge=ch, blacklist=["size", "gain", "feedback"])
+# create the optimizer
+gopt = ev.GeneticOptimizer( ev.fitness_in_environment, ev.compare_fitness, env=env )
+
+# initialize the population by using the
+N = 100 # use 100 for final run
+p = 0.1
+k = 2
+ertop = ev.ErdösRenyiTopology( p )
+ringtop = ev.RingTopology( k )
+# choose gain and feedback as in the paper
+gener  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop], p = [1.0, 0]), gain = 1.2, feedback = 2 )
+genrg  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop], p = [0, 1.0]), gain = 1.2, feedback = 2 )
+ev.init_population!(gopt, [(gener, 0.5), (genrg, 0.5)], 10)
+
+if N != 100
+  info("this seems to be a test run. use N=100 to create final results")
+end
+
+for i = 1:50
+  ev.step!(gopt)
+  ev.save_evolution("ring_vs_er.dat", gopt)
+end
+

### 2015-08-05 src/param/meta_parametric_object.jl

@@ -12,9 +12,15 @@ end
 
 # Meta Generator relative generator contribution
 function random_param( param::MetaCombinationParam, rng::AbstractRNG; s::Real = 0.1 )
+  @assert(length(param.val) != 1)
+
   # choose two paramerters
-  p1 = rand(rng, 1:length(param.val))
-  p2 = rand(rng, 1:length(param.val))
+  p1 = 0
+  p2 = 0
+  while p1 == p2
+    p1 = rand(rng, 1:length(param.val))
+    p2 = rand(rng, 1:length(param.val))
+  end
 
   v1 = param.val[p1]
   v2 = param.val[p2]
@@ -26,15 +32,20 @@ function random_param( param::MetaCombinationParam, rng::AbstractRNG; s::Real = 
   # select a change rate
   change = min(v1, 1-v2, abs(randn(rng) * s))
 
-  v1 += change
-  v2 -= change
+  v1 -= change
+  v2 += change
 
   # scale back
   v1 *= sum
   v2 *= sum
 
-  param.val[p1] = v1
-  param.val[p2] = v2
+  # new param
+  new_value = deepcopy(param)
+
+  new_value.val[p1] = v1
+  new_value.val[p2] = v2
+
+  return new_value
 end
 
 #####################################################################################

### 2015-08-05 research/er_ring.jl
bigger sim

@@ -21,13 +21,13 @@ ringtop = ev.RingTopology( k )
 # choose gain and feedback as in the paper
 gener  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop], p = [1.0, 0]), gain = 1.2, feedback = 2 )
 genrg  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop], p = [0, 1.0]), gain = 1.2, feedback = 2 )
-ev.init_population!(gopt, [(gener, 0.5), (genrg, 0.5)], 10)
+ev.init_population!(gopt, [(gener, 0.5), (genrg, 0.5)], 100)
 
 if N != 100
   info("this seems to be a test run. use N=100 to create final results")
 end
 
-for i = 1:50
+for i = 1:100
   ev.step!(gopt)
   ev.save_evolution("ring_vs_er.dat", gopt)
 end

### 2015-08-06 paper/paper.tex
added second picture
first text for result

@@
     \newcommand{\RINGk}{k}
+    \newcommand{\ERtoRING}{r}
     \newcommand{\FFlayers}{l}
@@
 \paragraph{Genetic Optimization}
-Due to the stochastic nature of the fitness function $F_C(G)$ of a genotype
-$G$, a very high number of samples would be required to obtain sufficiently
+Due to the stochastic nature of a genotype $G$'s fitness $F_C(G)$,
+a very high number of samples would be required to obtain sufficiently
 smooth approximations of $F_C(G)$ for typical optimization methods like
@@
 feasible.
-
-%and the fact that single evaluations of $F_C$ are
-%computationally very expensive, thus preventing increasing the number of
-%samples such that the estimation $E(F_C)$ becomes sufficiently smooth,
-%typical optimization methods like gradient descent or simulated annealing
-%are not usable.
 
@@
 
-\paragraph{Choice of Parameters}
-- $N=100$ fixed, since greater $N$ always increased quality and computational cost
-increased quadraticaly to $N$
-- Mutation rate
+\paragraph{Choice of Parameters and Algorithmic Quantities}
+For all of the simulations we chose the size of the networks to be fixed
+with $N=100$. Though increasing $N$ generally improved the quality of the
+learning results, the computational cost increased proportional to $N^2$,
+which made using networks as large as in the original FORCE publication
+(between $500$ and $20000$) utterly unusable for our purposes.
+
+The mutation rate of the parameters $\Theta$ of the genotypes\dots
 
 \section{Results}
-\subsection{Frequency Dependent Optimizations}
-- short description: get feeling for frequency range (by using unoptimized parameters
-from the force paper), make optimizations with 3 different challenges: 1) challenges that cover the frequency area completely, challenges that cover low frequencies, and challanges that cover high frequencies, then pick genotypes with high fitness.
-\subsection{Topology Optimizations}
-Baseline
-gain and feedback tests
+\subsection{Optimizing Genotypes For Frequency Ranges}
+Since we wanted to examine how the ideal genotypes vary for different
+types of problems, we examined how the choice of a frequency range affects
+the optimized genotype parameters.
+To this end we looked at sparse random networks (Erdös-Renyi topology with
+connection probability $\ERprob$) and conducted optimizations of the
+parameters $\gain$, $\feedback$ and $\ERprob$ for both high ($C\ix{high}$)
+and low ($C\ix{low}$) frequency challenges (see FIG.
+\ref{fig:freq_optimization}).
 
-(feedback shifts possible frequencies to higher values)
+\begin{figure}[hbct]
+    {\footnotesize\input{pictures/freq_optimization}}
+    \caption{
+        Genetic optimization process of genotypes with Erdös-Renyi topology
+        ($N=100$ fixed) for the challenges $C\ix{high}$ and $C\ix{low}$.
+        In each figure, one can see the $50$ genotypes of every Generation
+        (single dots) together with the respective mean values (thick
+        lines). Figures (a) to (c) depict the evolution of the three
+        parameters ($\gain$, $\feedback$, and $\ERprob$) being optimized,
+        while figure (d) shows the development of the fitness values.
+    }
+    \label{fig:freq_optimization}
+\end{figure}
 
-success of omega: erdös-renyi, ring
-optimizations: ER topology, Ring Topology, Mixing
+The most striking result is the strong growth of the feedback for
+$C\ix{high}$ to values higher than 14, while the value for $C\ix{low}$
+converges quickly to approximately 1 (FIG. \ref{fig:freq_optimization} (b)).
+This behavior may indicate, that the feedback pathway must
+overshadow the (maybe too slow) internal mechanics for high frequencies,
+while it is the other way round for low frequencies.
+Indeed, FIG. \ref{fig:freq_optimization} (a) shows that genotypes with
+a lower gain value (on average) are more successful for high frequencies,
+which seems to fit in: excessive recurrent dynamics with a long time
+scale (compared to the period to be learned) may disturb the learning.
+That the single genotypes remain quite scattered throughout the whole
+optimization process, though, implies that the exact gain value is not
+critical for the success of a genotype.
+
+Another interesting aspect is the branching process of $p$ that takes place
+for $C\ix{low}$ in FIG. \ref{fig:freq_optimization} (c). Albeit the branch
+seems to be dying out towards the end, it suggests, that very high $p$
+values don't ruin the networks capacity to learn low frequencies. This
+behavior was observed neither for high frequencies nor for optimizations
+over the full frequency regime.
+
+But how well did the fitness actually improve during the optimization?
+When looking at FIG. \ref{fig:freq_optimization} (d), one sees that the
+fitness for $C\ix{high}$ increased during nearly the whole process, while
+the algorithm failed to further optimize the genotypes' fitness for
+$C\ix{low}$ after about 20 generations. A major reason for this is
+probably, that the initial population for the $C\ix{low}$ optimization
+already contained genotypes with parameters similar to the (at least
+locally) optimal ones.
+
+One can also look at the success of the optimization algorithm in a more
+specific way: FIG. \ref{fig:freqrange} shows how well an unoptimized
+genotype $G_0$ handles fixed-frequency-challenges when compared to the
+genotypes $G\ix{high}$ and $G\ix{low}$ optimized for
+$C\ix{high}$ and $C\ix{low}$. Here it becomes clear that $G\ix{high}$ can
+indeed succeed at higher frequencies than $G_0$, but that it also fails at
+learning lower frequencies -- which was expected. On the contrary, while
+the high frequency capacities of $G\ix{low}$ have decreased when compared
+to $G_0$, the low frequency capacities have only marginally improved. This
+reinforces the impression that the optimization was unable to produce
+genotypes that can cope with low frequencies.
 
@@
     {\footnotesize\input{pictures/freqrange}}
-    \caption{CAPTION TEXT}
+    \caption{
+        The covered frequency range for an unoptimized
+        genotype $G_0$ as well as genotypes $G\ix{high}$ and $G\ix{low}$
+        that are optimized for $C\ix{high}$ respectively $C\ix{low}$. Each
+        of the three genotypes was tested against constant challenges
+        $C_\omega$ that yielded simple sinusoidal waves with frequency
+        $\omega$ for every seed value. The topology of all genotypes was
+        a mere Erdös-Renyi topology.
+        The parameters for the unoptimized genotype $G_0$ were taken to be
+        default values from \cite{FORCE} (which also uses Erdös-Renyi
+        topologies), while we chose genotypes $G\ix{high}$ and $G\ix{low}$
+        that did particularly well when explicitly optimizing for high and
+        low frequencies (see FIG. \ref{fig:freq_optimization} for more
+        details).
+    }
     \label{fig:freqrange}
 \end{figure}
+
+\subsection{Topology Optimizations}
+To find out whether the topology can be used to improve the learning, we
+performed an optimization in which the networks were an interpolation
+between sparse random connections (Erdös-Renyi) and a ring topology where
+the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
+fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
+Since we are interested in the influence of the topology, the parameters
+gain and feedback are kept constant during the optimization process,
+leaving the connection probability for ER topology $\ERprob$, the ring
+connection length $\RINGk$ and the interpolation value $\ERtoRING$ between
+ER and ring as free parameters.
+
+The results of the optimization are depicted in FIG. \ref{}. Unfortunately,
+the improvement in fitness after the first few generations in extremely
+low. The development of the ration parameter, however, unambiguously shows
+that the selection process prefers generators that consist mostly of ring
+topology with ring parameter $\RINGk \approx 1$. The ER connection probability
+shows a very broad distribution, which is simply due to the low dependence
+of the generated networks on $\ERprob$ when the usage of ER topology
+$\ERtoRING$ approaches zero.
 
@@
 - Problems
-    * Many magic parameters scattered in the model that were not examined exhaustively
+    * Many magic / ``empiric'' parameters scattered in the model that were not examined exhaustively
     * Only few and uncomprehensive number of test / experiments conducted due to the
@@
       of challenges than parameters (partially) -> must be very careful how to
-      interpret the results
-
-\blindtext
+      interpret the results / should have used fixed array of challenges instead
+      of random sampling
 

### 2015-08-06 paper/paper.tex
text

--- 2015-08-05 paper/paper.tex
@@
 \subsection{Frequency Dependent Optimizations}
-- short description: get feeling for frequency range (by using unoptimized parameters
-from the force paper), make optimizations with 3 different challenges: 1) challenges that cover the frequency area completely, challenges that cover low frequencies, and challanges that cover high frequencies, then pick genotypes with high fitness.
+- short description: get feeling for frequency range (by using unoptimized
+parameters from the force paper), make optimizations with 3 different
+challenges: 1) challenges that cover the frequency area completely,
+challenges that cover low frequencies, and challenges that cover high
+frequencies, then pick genotypes with high fitness.
 \subsection{Topology Optimizations}
@@
     {\footnotesize\input{pictures/freqrange}}
+    \caption{CAPTION TEXT}
+    \label{fig:freqrange}
+\end{figure}
+
+\begin{figure}
+    {\footnotesize\input{pictures/freq_optimization}}
     \caption{CAPTION TEXT}
@@
 
-\blindtext
+Our investigation showed that the frequency of the wave that is to be reconstructed 
+strongly determines whether FORCE learning is able to learn that curve. Still, 
+a good choice of network parameters is still able to improve the success rate for
+certain frequency ranges. Some parameters ($p$, $feedback$) can vary in a relatively large interval
+without influencing the success very much, whereas others are confined to a relatively small interval (gain).
+
+The code developed for this paper provides a framework which can be used to perform a variety of similar 
+investigations, e.g. the inclusion of feed forward topologies, or the extension to task other than simple 
+sinusoidals. To that end, it would be wise to improve that selection algorithm by using the same tasks
+for all generators in each generation, instead of evaluating them all independently. This would make the 
+selection process more robust because it removes the dependence on the task distribution.
+
+One problem with the approach presented here is that, in order to use the results of the genetic algorithm, 
+we determined a single set of parameters by taking the mean of the values of the most successful generators, 
+any information about an interdependence between the parameters is lost. We checked that there are no strong
+correlations between any pair of parameters by doing 3D plots that showed the two parameters for each generator 
+in development with the generations and could not detect any obvious correlation, but have not investigated 
+this problem thoroughly.
+
+The biggest problem with the current implementation, however, is the large amount of "magic" numbers, 
+i.e. hyper-parameters of the optimization, that are used in the code. Their values were determined
+by an informal search of testing a few optimizations. However, this should mostly affect the performance
+of the optimization algorithm and not the results themselves.
 

### 2015-08-06 paper/paper.tex
fits on 4 pages again

--- 2015-08-06 paper/paper.tex
@@
 \section{Introduction}
-Looking at the field of artificial neuronal networks, one finds a great
+Looking at the field of artificial neuronal networks (NN), one finds a great
 variety of different models describing both the dynamic of networks as well
@@
 Using the above questions as a guideline, we decided to analyze what
-network characteristics make the FORCE algorithm, introduced in
-\citep{FORCE}, especially suitable for learning periodic patterns.  More
-precisely, we wanted to identify which properties of a network are
+network characteristics make the FORCE algorithm \citep{FORCE} especially suitable
+for learning periodic patterns.  More
+precisely, we wanted to identify which properties of a NN are
 important for its learning success, and how to choose their
@@
 also wanted them to follow certain patterns and allow for specialization:
-How dense are the synapses distributed? Is there a strong compartmentalization?
-Are recurrent structures frequent? How strong should the feedback signal
+How densely are the synapses distributed? Is there a strong compartmentalization?
+How strong should the feedback signal
 be, and how strong should the synapses fire on average?
 
-To address these issues we looked at parameterized random networks, or in
-the language of genetics, at different \emph{network genotypes}. These
+To address these issues we looked at parameterized random networks (in
+the language of genetics, as we are using a genetic optimization algorithm, at different 
+\emph{network genotypes}). These
 genotypes contain parametric information about the networks
 (\emph{phenotypes}) to be created. 
-Different phenotypes with the same genotype may thus have strong
+Different phenotypes of the same genotype may thus have strong
 quantitative differences, but they share structural qualities whose fitness
@@
 
-Since calculating the fitness of a genotype required the evaluation of many
+Since calculating the fitness of a genotype requires the evaluation of many
 different phenotypes and because of the resulting stochastic nature of the
@@
 %partially abstract fashion in the next section.
-rather high-level overview. 
-More details on the actual implementation are then provided in the third
-section (or ultimately in the published source
-code / supplementary material).
-Finally, some of the obtained results are presented and discussed.
+rather high-level overview, followed by 
+more details on the actual implementation.
+Finally, the obtained results are presented and discussed.
 
@@
 We chose the same rate based network architecture
-that is used in the original FORCE publication \citep{FORCE} (denoted as
+that is used in the original FORCE publication \citep{FORCE} (therein denoted as
 architecture A) and also in \citep{RM}, where our notation mainly stems
@@
 $\omega\Ix{rec}_{ij}$ and the feedback synapse weights
-$\omega\Ix{fb}_{ij}$. One can see that the neurons exponentially
-decay if they are not presented stimulus by other neurons. 
+$\omega\Ix{fb}_{ij}$. 
 We solved this system of differential equations by
@@
 neurons should eventually match a given periodic target pattern $t\mapsto
-r'_i(t)$, called a \emph{task} from now on, as accurately as possible. To
+r'_i(t)$, henceforth called a \emph{task}, as accurately as possible. To
 update $\omega\Ix{read}$ we chose a variant of the FORCE algorithm
@@
 generator that returns a network $\mathcal{N}=G(\lambda)$ for every seed
-value $\lambda$.
+$\lambda \in \mathbb{N}$.
 The genotypes' parameters $\theta\in\Theta$ were either integer or floating
@@
 $\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
-occupation probability $p$ (for random Erdös-Renyi networks), the
-neighborhood range $k$ (for ring topologies), or the number $l$ of layers
-(for feed-forward networks). We furthermore introduced ratio parameters
+occupation probability $p$ (for random Erdös-Renyi networks), and the
+neighborhood range $k$ (for ring topologies). We furthermore introduced ratio parameters
 that allowed us to interpolate between different topologies. Optimized
@@
 \paragraph{Fitness} The next step was to define the \emph{fitness} of a given
-genotype when confronted with a diverse range of tasks, like waves of
+genotype when confronted with a diverse range of tasks, e.g. waves of
 different frequencies. To express this formally, we introduced
@@
 distribution. If we now make use of a success function $S(\mathcal{N},
-r')\in\{0, 1\}$ that decides, if the network $\mathcal{N}$
+r')\in\{0, 1\}$ that decides whether the network $\mathcal{N}$
 has successfully reproduced the desired output $r'$, the fitness $F_C(G)$
@@
 For the optimization to work properly, the tasks presented to the networks
-must be reasonably chosen; they must be learnable by FORCE (at least in
+must be chosen reasonably; they must be learnable by FORCE (at least in
 principle) and must be compatible with technical aspects like the
 integration time step. 
-%have (for the most part) to be solvable by FORCE. 
 While it was shown in \citep{FORCE} that large recurrent neural networks are
@@
 As mentioned in \emph{Concepts and Methods}, we need a function $S$
-that decides weather a network succeeded or failed to solve a task.
+that decides whether a network succeeded or failed to solve a task.
 To this end
@@
 
-Instead, a measure that is time-shift independent and maximized for functions of
+A measure that is time-shift independent and maximized for functions of
 identical shape is the maximum of the cross correlation 
@@
 for a longer time (order of chunk size), but since we are going to work
-with superpositions of sinusoidals of different frequencies, this drawback
+with superpositions of sinusoidals, this drawback
 is unproblematic.
@@
 Due to the stochastic nature of a genotype $G$'s fitness $F_C(G)$,
-a very high number of samples would be required to obtain sufficiently
+a very high number of samples would be required to numerically obtain sufficiently
 smooth approximations of $F_C(G)$ for typical optimization methods like
 gradient descend or simulated annealing. But as the computational cost for
-even one learning process was considerable high, these methods were not
+even one learning process is considerable high, these methods were not
 feasible.
 
-An alternative approach would be to use techniques from Monte Carlo
-integration, e.g.  stratified or importance sampling, to aid in the
-evaluation of $F_C(G)$, but since $G(\lambda)$ and $C(\lambda)$ depend
-non-continuously on the seed value $\lambda$, this too would not solve the
-problem.
+%An alternative approach would be to use techniques from Monte Carlo
+%integration, e.g.  stratified or importance sampling, to aid in the
+%evaluation of $F_C(G)$, but since $G(\lambda)$ and $C(\lambda)$ depend
+%non-continuously on the seed value $\lambda$, this too would not solve the
+%problem.
 
@@
  \item[Evaluation] The fitness $F_i = F_C(G_i)$ of each individual is
-     measured using at least 20 samples (and adaptively more if the
+     measured using at least 50 samples (and adaptively more if the
      $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
@@
 
-\item[Reproduction] We obtain the next generation $\mathcal{G}_{t+1}$ by first
-    generating $2N$ candidates via \emph{mutating} or \emph{recombining}
-    elements of $\mathcal{G}\Ix{survive}_t$. This means we either change
+\item[Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
+it into the next generation by eith \emph{mutating} or \emph{recombining} the 
+survivors and get $\mathcal{G}_{t+1}$. This means we either change
     a single parameter randomly or take two individuals and randomly choose
-    for each parameter from which parent to take the value.  We then
-    perform a single sample fitness test and remove all networks that fail,
-    until only $N$ networks are left; these form $\mathcal{G}_{t+1}$. By
-    this approach the most unpromising elements of the new generation are
-    removed before the expensive fitness evaluation takes place. 
+    for each parameter from which parent to take the value.   
 \end{description}

### 2015-08-06 paper/paper.tex

--- 2015-08-06 paper/paper.tex
@@
 
+Our investigation showed that the frequency of the wave that is to be reconstructed 
+strongly determines whether FORCE learning is able to learn that curve. Still, 
+a good choice of network parameters is still able to improve the success rate for
+certain frequency ranges. Some parameters ($p$, $feedback$) can vary in a relatively large interval
+without influencing the success very much, whereas others are confined to a relatively small interval (gain).
+
+The code developed for this paper provides a framework which can be used to perform a variety of similar 
+investigations, e.g. the inclusion of feed forward topologies, or the extension to task other than simple 
+sinusoidals. To that end, it would be wise to improve that selection algorithm by using the same tasks
+for all generators in each generation, instead of evaluating them all independently. This would make the 
+selection process more robust because it removes the dependence on the task distribution.
+
+One problem with the approach presented here is that, in order to use the results of the genetic algorithm, 
+we determined a single set of parameters by taking the mean of the values of the most successful generators, 
+any information about an interdependence between the parameters is lost. We checked that there are no strong
+correlations between any pair of parameters by doing 3D plots that showed the two parameters for each generator 
+in development with the generations and could not detect any obvious correlation, but have not investigated 
+this problem thoroughly.
+
+The biggest problem with the current implementation, however, is the large amount of "magic" numbers, 
+i.e. hyper-parameters of the optimization, that are used in the code. Their values were determined
+by an informal search of testing a few optimizations. However, this should mostly affect the performance
+of the optimization algorithm and not the results themselves.
+
 \bibliography{sources}

### 2015-08-06 paper/paper.tex
commented some stuff out

@@
 
-Since calculating the fitness of a genotype requires the evaluation of many
-different phenotypes and because of the resulting stochastic nature of the
-problems we faced, we decided to use genetic algorithms for the
-optimization process.
+% TODO we say the same thing later on again. maybe, the comment about the genetic algorithm
+% should be moved out of the parenthesis
+%Since calculating the fitness of a genotype requires the evaluation of many
+%different phenotypes and because of the resulting stochastic nature of the
+%problems we faced, we decided to use genetic algorithms for the
+%optimization process.
 
@@
 integration time step. 
-While it was shown in \citep{FORCE} that large recurrent neural networks are
-capable of learning very complex tasks, the optimization requires an
-extensive number of learning processes, since every single evaluation of
-the fitness $F_C$ needs dozens of tested phenotypes (see below). Therefore,
-the networks had to be restricted to being quite small ($\approx 100$
-neurons) and are consequently less powerful.
+%While it was shown in \citep{FORCE} that large recurrent neural networks are
+%capable of learning very complex tasks, the optimization requires an
+%extensive number of learning processes, since every single evaluation of
+%the fitness $F_C$ needs dozens of tested phenotypes (see below). Therefore,
+%the networks had to be restricted to being quite small ($\approx 100$
+%neurons) and are consequently less powerful.
 
@@
 possible frequencies, we choose the frequencies for the optimization tasks
-from the slightly larger interval $[0.6e-2, 4]$ such that their logarithms
+from the slightly larger interval $[\num{0.6e-2}, 4]$ such that their logarithms
+% TODO this number is no longer right: we have low, high and full
 are uniformly spaced. This allows for good resolution for both high and low
@@
 timescale of a single chunk (which is about 10 seconds of simulated time).
-This measure becomes problematic for functions that are very close to zero
-for a longer time (order of chunk size), but since we are going to work
-with superpositions of sinusoidals, this drawback
-is unproblematic.
+%This measure becomes problematic for functions that are very close to zero
+%for a longer time (order of chunk size), but since we are going to work
+%with superpositions of sinusoidals, this drawback
+%is unproblematic.
 
@@
      $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
- \item[Selection] All individuals are grouped into disjunct pairs for which
-     we check which one has the higher fitness. This is repeated $k=10$ times,
-     counting the number of wins for each individual. Based thereon we use the
-     $p=50\%$ best individuals $\mathcal{G}\Ix{survive}_t$ to build the next
-     generation.  By using pair comparisons instead of simple ordering
-     according to the fitness, we allow weaker individuals to reach
-     the next generation (albeit with low probability), keeping the gene
-     pool more diverse.
+ \item[Selection] We choose $p=50\%$ of the most successful members of $G_t$ as survivors.
+With a small chance, a weaker generator can survive, in order to keep the gene pool more diverse.
 
 \item[Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
-it into the next generation by eith \emph{mutating} or \emph{recombining} the 
+it into the next generation by either \emph{mutating} or \emph{recombining} the 
 survivors and get $\mathcal{G}_{t+1}$. This means we either change
@@
 - Problems
-    * Many magic / ``empiric'' parameters scattered in the model that were not examined exhaustively
     * Only few and uncomprehensive number of test / experiments conducted due to the
       computational cost.
-    * Test with several other topologies did not yield concrete results,
-      only rough tendencies.
     * Influence of statistics: Learning success depends stronger on choice

### 2015-08-06 research/q_of_omega_ring.jl
q of omega for ring

@@ -1,0 +1,24 @@
+push!(LOAD_PATH, "/scratch01/eriks/neural/Neuronal-Plasticity-Seminar/src/")
+
+using EvoNet
+#=ev = EvoNet=#
+
+# F_C(ω) curve for an unoptimized genotype
+
+@everywhere function rating(f)
+  N = 100 # use 100 for final run
+  p = .07
+  k = 1.9
+  ertop = EvoNet.ErdösRenyiTopology(p)
+  ringtop = ev.RingTopology( k )
+  
+  gen  = ev.SparseFRGenerator( N, topology = ev.MetaTopology([ertop, ringtop], p = [0.02, 0.98]), gain = 1.2, feedback = 2 )
+  ch   = EvoNet.simple_wave(amplitude=1, frequency=f, offset=0)
+  env  = EvoNet.Environment(challenge=ch)
+  s = EvoNet.fitness_in_environment(gen, samples=200, env=env)
+  return EvoNet.get_value(s)
+end
+
+freq = logspace(-3, 0.5, 1000)
+result = pmap(x->rating(x), freq)
+writedlm("freqrange_ring.txt", [freq result])

### 2015-08-06 paper/paper.tex
introduction adapted

@@
     \usepackage{enumitem}
+    \usepackage{siunitx}
 
@@
 as their plasticities, i.e.\ their learning behavior. 
-But what exactly makes artificial neuronal networks learn successfully? And
-why are certain types of networks and certain learning rules effective for
-some problems, but fail at solving others? Confronting and eventually
-answering these questions is critical when trying to gain a deep
-understanding of neural networks and artificial intelligence.
+But, given some dynamics and plasticity rules, which features of a network
+determine, if it is a good candidate for doing a specific task?
+%But what exactly makes artificial neuronal networks learn successfully? And
+%why are certain types of networks and certain learning rules effective for
+%some problems, but fail at solving others? 
+%Confronting and eventually
+%answering these questions is critical when trying to gain a deep
+%understanding of neural networks and artificial intelligence.
 
-Using the above questions as a guideline, we decided to analyze what
-network characteristics make the FORCE algorithm \citep{FORCE} especially suitable
-for learning periodic patterns.  More
-precisely, we wanted to identify which properties of a NN are
-important for its learning success, and how to choose their
-values for an optimal result.  To this end, one option would be to take an
-initial network and an adequate periodic pattern to be learned, and
-optimize the network's internals for this very task. Then patterns may
-emerge that hint at how an optimal network design looks.
+Using this question as a guideline, we decided to examine network
+characteristics that make the FORCE algorithm \citep{FORCE} especially
+suitable for learning periodic patterns. More precisely, we wanted to
+identify which high level properties of a NN are important for its learning
+success under FORCE, and how to choose their values for an optimal result.
 
-However, we took a different approach:
-Instead of fine tuning single internal weights, we were concerned with the
-general rules of the network's structure. So we wanted our networks to
-be random to a certain degree, which seems biologically plausible, but we
-also wanted them to follow certain patterns and allow for specialization:
-How densely are the synapses distributed? Is there a strong compartmentalization?
-How strong should the feedback signal
-be, and how strong should the synapses fire on average?
+%To this end, one option would be to take an
+%initial network and an adequate periodic pattern to be learned, and
+%optimize the network's internals for this very task. Then patterns may
+%emerge that hint at how an optimal network design looks. 
+%However, we took a different approach:
+Instead of fine tuning single internal weights, we were thus concerned with the
+general rules of the network's properties. And we wanted our networks to
+be random to a certain degree, which seems biologically plausible; but we
+also wanted to allow for structural specialization:
+How densely are the synapses distributed? How strong shall the feedback
+signal be, and how strong should the synapses fire on average?
+%To address these issues we looked at parameterized random networks (in
+%the language of genetics, as we are using a genetic optimization algorithm, at different 
+%\emph{network genotypes}).
 
-To address these issues we looked at parameterized random networks (in
-the language of genetics, as we are using a genetic optimization algorithm, at different 
-\emph{network genotypes}). These
-genotypes contain parametric information about the networks
-(\emph{phenotypes}) to be created. 
-Different phenotypes of the same genotype may thus have strong
-quantitative differences, but they share structural qualities whose fitness
-for a preassigned range of tasks we wanted to analyze and improve.
-
+To address these issues we introduce \emph{network genotypes}, which
+contain parametric information about the networks (\emph{phenotypes}) to be
+created: Different phenotypes of the same genotype may have strong
+differences in detail, but they share structural qualities of interest.
+Improving these genotypes for a specific range of tasks via genetic
+optimization is the main concern of this article.
+%The fitness of these genotypes for a specific range of tasks was optimized 
+%using genetic algorithms.
+%whose fitness
+%for a preassigned range of tasks we analyze and improve with genetic
+%optimization.
 % TODO we say the same thing later on again. maybe, the comment about the genetic algorithm
@@
 %partially abstract fashion in the next section.
-rather high-level overview, followed by 
-more details on the actual implementation.
-Finally, the obtained results are presented and discussed.
+rather high-level overview, followed by more details on the actual
+implementation.  Finally, the obtained results are presented and discussed.
 

### 2015-08-06 paper/paper.tex
introduction improved

@@
 But, given some dynamics and plasticity rules, which features of a network
-determine, if it is a good candidate for doing a specific task?
+determine whether it is a good candidate for doing a specific task?
 %But what exactly makes artificial neuronal networks learn successfully? And
@@
 
-Using this question as a guideline, we decided to examine network
-characteristics that make the FORCE algorithm \citep{FORCE} especially
-suitable for learning periodic patterns. More precisely, we wanted to
-identify which high level properties of a NN are important for its learning
-success under FORCE, and how to choose their values for an optimal result.
+%Using this question as a guideline, we decided to examine network
+%characteristics that make the FORCE algorithm \citep{FORCE} especially
+%suitable for learning periodic patterns. More precisely, we wanted to
+%identify which high level properties of a NN are important for its learning
+%success under FORCE, and how to choose their values for an optimal result.
+
+Using this question as a guideline, we decided to analyze which high level
+properties of a NN are important for its learning success under a variant
+of the FORCE algorithm \citep[Architecture A]{FORCE}, and how to choose
+their values for an optimal result.
 
@@
 
-To address these issues we introduce \emph{network genotypes}, which
-contain parametric information about the networks (\emph{phenotypes}) to be
-created: Different phenotypes of the same genotype may have strong
-differences in detail, but they share structural qualities of interest.
-Improving these genotypes for a specific range of tasks via genetic
-optimization is the main concern of this article.
+To address these issues in the context of genetic algorithms, we introduce
+\emph{network genotypes}, which contain parametric information about the
+networks (\emph{phenotypes}) to be created: Different phenotypes of the
+same genotype may have strong differences in detail, but they share
+structural qualities of interest.  Improving these genotypes for a specific
+range of tasks via genetic optimization is the main concern of this
+article.
 %The fitness of these genotypes for a specific range of tasks was optimized 

### 2015-08-06 paper/paper.tex
some text cuts

@@
 by \footnote{~Note that we did not include external input signals in this
-formula, as is done in \citep{FORCE} and \citep{RM}. We in fact enabled
-inputs in our code base, but did not use them for the main results.}
+formula, as is done in \citep{FORCE} and \citep{RM}.}
 \begin{equation}
@@
 We solved this system of differential equations by
-a simple Newton method with integration time steps $dt = 0.01$, which was
+Euler integration with a time steps $dt = 0.01$, which was
 also used in \citep{FORCE}. 
@@
 r'_i(t)$, henceforth called a \emph{task}, as accurately as possible. To
-update $\omega\Ix{read}$ we chose a variant of the FORCE algorithm
-\citep{FORCE} that rapidly modifies the feedback loop to keep the errors
+update $\omega\Ix{read}$ we chose variant A of the FORCE algorithm
+\citep{FORCE}, which rapidly modifies the feedback loop to keep the errors
 $|r_i(t) - r'_i(t)|$ small from the beginning on.
@@
 While the authors of \citep{FORCE} update the weights $\lambda\Ix{read}$ in
-fixed intervals (every second time step), we used an adaptive mechanism to
-determine how often learning should occur. This was initially introduced
-for performance reasons, as the single learning steps proved to be the main
-computational bottleneck, but we also found that the long time network
-dynamics tended to be slightly more accurate for adaptive learning steps.
+fixed intervals, we used an adaptive mechanism to
+determine how often learning should occur. 
+%This was initially introduced
+%for performance reasons, as the single learning steps proved to be the main
+%computational bottleneck, but we also found that the long time network
+%dynamics tended to be slightly more accurate for adaptive learning steps.
 
 \paragraph{Network Genotypes} Since we are not
-interested in single network properties but rather want to find out which
+interested in single network properties, but rather want to find out which
 structural elements of networks are important for learning success, we
@@
 $\lambda \in \mathbb{N}$.
-The genotypes' parameters $\theta\in\Theta$ were either integer or floating
+The genotypes' parameters $\theta\in\Theta$ are either integer or floating
 point values and can be classified as affecting either (1) the topology of
@@
 occupation probability $p$ (for random Erdös-Renyi networks), and the
-neighborhood range $k$ (for ring topologies). We furthermore introduced ratio parameters
+neighborhood range $k$ (for ring topologies). We furthermore introduced ratio parameters $r$
 that allowed us to interpolate between different topologies. Optimized
@@
 distribution. If we now make use of a success function $S(\mathcal{N},
-r')\in\{0, 1\}$ that decides whether the network $\mathcal{N}$
+r')\in\{0, 1\}$ that decides whether the NN $\mathcal{N}$
 has successfully reproduced the desired output $r'$, the fitness $F_C(G)$
@@
 \paragraph{Choice of Parameters and Algorithmic Quantities}
-For all of the simulations we chose the size of the networks to be fixed
-with $N=100$. Though increasing $N$ generally improved the quality of the
-learning results, the computational cost increased proportional to $N^2$,
-which made using networks as large as in the original FORCE publication
-(between $500$ and $20000$) utterly unusable for our purposes.
-
 The mutation rate of the parameters $\Theta$ of the genotypes\dots

### 2015-08-06 paper/paper.tex
small fixes

@@
 converges quickly to approximately 1 (FIG. \ref{fig:freq_optimization} (b)).
-This behavior may indicate, that the feedback pathway must
+This behavior may indicate that the feedback pathway must
 overshadow the (maybe too slow) internal mechanics for high frequencies,
@@
 specific way: FIG. \ref{fig:freqrange} shows how well an unoptimized
-genotype $G_0$ handles fixed-frequency-challenges when compared to the
+genotype $G_0$ handles fixed-frequency challenges when compared to the
 genotypes $G\ix{high}$ and $G\ix{low}$ optimized for

### 2015-08-07 paper/paper.tex
paper updated

@@
 \section{Introduction}
-Looking at the field of artificial neuronal networks (NN), one finds a great
+Looking at the field of artificial neuronal networks, one finds a great
 variety of different models describing both the dynamic of networks as well
@@
 Using this question as a guideline, we decided to analyze which high level
-properties of a NN are important for its learning success under a variant
-of the FORCE algorithm \citep[Architecture A]{FORCE}, and how to choose
-their values for an optimal result.
+properties of a neuronal network are important for its learning success
+under a variant of the FORCE algorithm \citep[Architecture A]{FORCE}, and
+how to choose their values for an optimal result.
 
@@
 %However, we took a different approach:
-Instead of fine tuning single internal weights, we were thus concerned with the
-general rules of the network's properties. And we wanted our networks to
-be random to a certain degree, which seems biologically plausible; but we
-also wanted to allow for structural specialization:
-How densely are the synapses distributed? How strong shall the feedback
-signal be, and how strong should the synapses fire on average?
+Instead of fine tuning every single internal weight, we were thus concerned
+with the general rules of the network's properties. And we wanted our
+networks to be random to a certain degree, which seems biologically
+plausible; but we also wanted to allow for structural specialization: How
+densely are the synapses distributed? How strong shall the feedback signal
+be, and how strong should the synapses fire on average?
 %To address these issues we looked at parameterized random networks (in
@@
 \paragraph{Network Model}
-%MAYBE WE SHOULD MOTIVATE FORCE HERE, WHICH THEN NATURALLY LEADS TO THE CHOICE OF NETWORK BELOW.
 
-We chose the same rate based network architecture
-that is used in the original FORCE publication \citep{FORCE} (therein denoted as
-architecture A) and also in \citep{RM}, where our notation mainly stems
-from. So we look at a network $\mathcal{N}$ with $N$ internal neurons and
-states $x_i$ for $1 \le i \le N$ that loosely represent the neurons'
-membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
-given by $r_i = \tanh x_i$.  Furthermore, the network contains $R$ readout
-neurons with states $z_i$ for $1 \le i \le R$, where
+We chose the same rate based network architecture that is used in the
+original FORCE publication \citep{FORCE} (therein denoted as architecture
+A) and also in \citep{RM}, where our notation mainly stems from. So we look
+at a network $\mathcal{N}$ with $N$ internal neurons and states $x_i$ for
+$1 \le i \le N$ that loosely represent the neurons' membrane potential. The
+firing rate $r_i$ of the $i$-th internal neuron is given by $r_i = \tanh
+x_i$.  Furthermore, the network contains a readout neuron with the state
 \begin{equation} 
-    z_i = \sum_{j=1}^{N}\omega\Ix{read}_{ij}\,r_j
+z = \sum_{j=1}^{N}\omega\Ix{read}_{j}\,r_j
 \end{equation}
-with the readout weights $\omega\Ix{read}$.
+where $\omega\Ix{read}$ denotes the readout weight vector.
 Taking into account both internal dynamics as well as an external feedback
@@
 where we introduce the internal recurrent synapse weights
-$\omega\Ix{rec}_{ij}$ and the feedback synapse weights
-$\omega\Ix{fb}_{ij}$. 
+$\omega\Ix{rec}$ and the feedback synapse weights
+$\omega\Ix{fb}$. 
 We solved this system of differential equations by
-Euler integration with a time steps $dt = 0.01$, which was
+Euler integration with a time steps $\mathrm{d}t = 0.01$, which was
 also used in \citep{FORCE}. 
@@
 \paragraph{Learning} Learning took place by stepwisely adapting the weights
-$\omega\Ix{read}$ so that the dynamics $t\mapsto r_i(t)$ of the readout
-neurons should eventually match a given periodic target pattern $t\mapsto
-r'_i(t)$, henceforth called a \emph{task}, as accurately as possible. To
+$\omega\Ix{read}$ in such a way that the readout dynamic $t\mapsto z(t)$
+should eventually match a periodic target pattern $t\mapsto
+z'(t)$, henceforth called a \emph{task}, as accurately as possible. To
 update $\omega\Ix{read}$ we chose variant A of the FORCE algorithm
 \citep{FORCE}, which rapidly modifies the feedback loop to keep the errors
-$|r_i(t) - r'_i(t)|$ small from the beginning on.
+$|z(t) - z'(t)|$ small from the beginning on.
 
 While the authors of \citep{FORCE} update the weights $\lambda\Ix{read}$ in
-fixed intervals, we used an adaptive mechanism to
-determine how often learning should occur. 
+fixed intervals, we used an adaptive mechanism to determine how often
+learning should occur. 
 %This was initially introduced
@@
 
-\paragraph{Network Genotypes} Since we are not
-interested in single network properties, but rather want to find out which
-structural elements of networks are important for learning success, we
-introduced  \emph{network genotypes}.
-Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
-which concrete networks are created when $G$ is expressed in a single
-phenotype. Technically $G$ may be understood as a stochastic network
-generator that returns a network $\mathcal{N}=G(\lambda)$ for every seed
-$\lambda \in \mathbb{N}$.
+\paragraph{Network Genotypes} 
+In order to find out which structural elements of networks are important
+for learning sucess, we introduce \emph{network genotypes}.
+%Since we are not interested in single network
+%properties, but rather want to find out which structural elements of
+%networks are important for learning success, we introduced  \emph{network
+%genotypes}.
+%Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
+%which concrete networks are created when $G$ is expressed in a single
+%phenotype. 
+Technically, a genotype $G$ with parameters $\Theta$ may be understood as
+a stochastic network generator that returns a network
+$\mathcal{N}=G(\lambda)$ for every seed value $\lambda$.
+
 The genotypes' parameters $\theta\in\Theta$ are either integer or floating
-point values and can be classified as affecting either (1) the topology of
+point values and can be classified as affecting (1) the topology of
 $\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
-$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
-occupation probability $p$ (for random Erdös-Renyi networks), and the
-neighborhood range $k$ (for ring topologies). We furthermore introduced ratio parameters $r$
-that allowed us to interpolate between different topologies. Optimized
-parameters of type (2) were the feedback strength $\feedback$ (with
-$\omega\Ix{fb}\propto \feedback$) and the gain value $\gain$ (with
-$\omega\Ix{rec} \propto \gain$).
+$\omega\Ix{fb}$. Optimized parameters of type (1) are the
+occupation probability $p$ for sparse random topologies (Erdös-Renyi
+topology), the neighborhood range $k$ for ring topologies, and a ratio
+parameter $q$ used to interpolate between Erdös-Renyi and ring
+topologies. Optimized parameters of type (2) are the feedback strength
+$\feedback\propto \omega\Ix{fb}$ and the gain value
+$\gain\propto\omega\Ix{rec}$.
 
-\paragraph{Fitness} The next step was to define the \emph{fitness} of a given
-genotype when confronted with a diverse range of tasks, e.g. waves of
-different frequencies. To express this formally, we introduced
+\paragraph{Fitness} The next step was to define the \emph{fitness} of
+a given genotype when confronted with a range of different tasks, e.g.
+waves of different frequencies. To express this formally, we introduce
 \emph{challenges}: For a given seed value $\lambda$, the challenge $C$
-yields a task $r' = C(\lambda)$ according to some (parametrized)
-distribution. If we now make use of a success function $S(\mathcal{N},
-r')\in\{0, 1\}$ that decides whether the NN $\mathcal{N}$
-has successfully reproduced the desired output $r'$, the fitness $F_C(G)$
-of a genotype $G$ for a challenge $C$ may reasonably be defined as the
-expectation value of $\lambda \mapsto S(G(\lambda), C(\lambda))$. For fixed
-challenges, the main task of optimizing the genotype is now reduced to
-finding parameters $\Theta$ that maximize $F_C$.
+yields a task $z' = C(\lambda)$. If we now apply a success function
+$S(\mathcal{N}, z')\in\{0, 1\}$ that decides whether the network $\mathcal{N}$
+has successfully reproduced the desired output $z'$, the fitness $F_C(G)$
+of a genotype $G$ for $C$ may be defined as the expectation value of
+$\lambda \mapsto S(G(\lambda), C(\lambda))$. For fixed challenges, the
+main problem of optimizing the genotype thus is to find parameters
+$\Theta$ that maximize $F_C$.
 
-% IDEA: Maybe put this part at the end of the paper. Most things needed to
-% understand the results (in principle) have already been mentioned at this
-% point.
 \section{Methods and Implementation}
@@
 
-\paragraph{Tasks}
-For the optimization to work properly, the tasks presented to the networks
-must be chosen reasonably; they must be learnable by FORCE (at least in
-principle) and must be compatible with technical aspects like the
-integration time step. 
-%While it was shown in \citep{FORCE} that large recurrent neural networks are
-%capable of learning very complex tasks, the optimization requires an
-%extensive number of learning processes, since every single evaluation of
-%the fitness $F_C$ needs dozens of tested phenotypes (see below). Therefore,
-%the networks had to be restricted to being quite small ($\approx 100$
-%neurons) and are consequently less powerful.
+\paragraph{High and Low Frequency Challenges}
+%For the optimization to work properly, the tasks presented to the networks
+%must be chosen reasonably; they must be learnable by FORCE (at least in
+%principle) and must be compatible with technical aspects like the
+%integration time step.
 
-To get a first estimate whether it might be possible to learn a given task
-using FORCE with 100 neurons in our network model, we looked at a single
-sinusoidal of frequency $\omega$ and determined the frequency range in
-which an (unoptimized) network (i.e. \gain, \feedback, \ERprob) is able to
-learn the function (see fig. \ref{}).
+Our concrete goal for this article was to examine which choice of genotype
+parameters are optimal for learning sinusoidal waves with relatively high
+or low frequencies. 
+Based on preliminary simulation results we chose
+$\nu\in[]$ as frequency range for the challenge $C\ix{high}$, producing
+high frequency waves, and $\nu\in[]$ for $C\ix{low}$, producing low
+frequency waves. 
+Optimizing the fitness functions $F_{C\ix{high}}$ or $F_{C\ix{low}}$ thus
+results in genotypes with the desired frequency properties.
 
-Since we expect/hope that the optimized network extends this range of
-possible frequencies, we choose the frequencies for the optimization tasks
-from the slightly larger interval $[\num{0.6e-2}, 4]$ such that their logarithms
-% TODO this number is no longer right: we have low, high and full
-are uniformly spaced. This allows for good resolution for both high and low
-frequencies.
+%Based on our simulation results we found that can be expected to be
+%learned using simple sinusoidal waves.
+%We therefore used challenges $C\ix{high}$ and $C\ix{low}$ to specifically
+%optimize genotypes for the high respectively low frequency region.
+
+%To get a first estimate whether it might be possible to learn a given task
+%using FORCE with 100 neurons in our network model, we looked at a single
+%sinusoidal of frequency $\omega$ and determined the frequency range in
+%which an (unoptimized) network (i.e. \gain, \feedback, \ERprob) is able to
+%learn the function (see fig. \ref{}).
+
+%Since we expect/hope that the optimized network extends this range of
+%possible frequencies, we choose the frequencies for the optimization tasks
+%from the slightly larger interval $[\num{0.6e-2}, 4]$ such that their logarithms
+%are uniformly spaced. This allows for good resolution for both high and low
+%frequencies.
 
 \paragraph{Success}
-As mentioned in \emph{Concepts and Methods}, we need a function $S$
-that decides whether a network succeeded or failed to solve a task.
-To this end
-%In order to optimize the function reconstruction of a force network,
-a measure $f(r', r)$ for the concordance of the target function $r'(t)$ and
-the network readout $r(t)$ is sought. Though one could use a simple
-difference based metric, like the norm $f\ix{2}(r, r')
-= \left\|r-r'\right\|_2$, this has the drawback that small phase
-mismatches could produce high differences, even though the
-shape of $r$ matches the one of $r'$ well.
+To actually compute the fitness, we need a function $S$ that decides
+whether a network succeeded to solve a task or not. To this end,
+a measure $f(z', z)$ for the concordance of the target function $t\to z'(t)$ and
+the network readout $t\to z(t)$ is sought. 
+Though one could use a simple difference based metric, this has the
+drawback that plain phase mismatches could produce high
+differences for properly matching shapes.
+
+%Though one could use a simple difference based metric, like the norm
+%$f\ix{2}(z, z') = \left\|r-r'\right\|_2$, this has the drawback that small
+%phase mismatches could produce high differences, even though the shape of
+%$r$ matches the one of $r'$ well.
 
 A measure that is time-shift independent and maximized for functions of
-identical shape is the maximum of the cross correlation 
+identical shape, on the contrary, is the maximum of the cross correlation 
 \begin{align}
-    f\ix{cor}(r, r') = \max_t \frac{\left<r, r'(\cdot
-    - t)\right>}{\sqrt{\left< r, r\right> \left<r', r' \right>}}~.
+    f\ix{cor}(z, z') = \max_t \frac{\left<z, z'(\cdot
+    - t)\right>}{\sqrt{\left< z, z\right> \left<z', z' \right>}}~.
 \end{align}
-Algorithmically, the signals $r$ and $r'$ are split into overlapping chunks
+Algorithmically, the signals $z$ and $z'$ are split into overlapping chunks
 which are correlated separately.  This saves computational power and
@@
 Using this measure of quality, the success function for a network
-$\mathcal{N}$ with readout $r$ for the task $r'$ was taken to be 
+$\mathcal{N}$ with readout $z$ for the task $z'$ was (by empirical means)
+taken to be 
 \begin{equation}
-    S(\mathcal{N}, r') = \begin{cases} 1\quad\mathrm{if}~f\ix{cor}(r, r') \ge 0.95,\\
-                                       0\quad\mathrm{if}~f\ix{cor}(r, r') < 0.95~.
+    S(\mathcal{N}, z') = \begin{cases} 1,\quad\mathrm{if}~f\ix{cor}(z, z') \ge 0.95,\\
+                                       0,\quad\mathrm{if}~f\ix{cor}(z, z') < 0.95~.
                          \end{cases}
 \end{equation}
-The threshold value of 0.95 was determined empirically, as signals $r$ with
-$f(r, r') \ge 0.95$ met our assessment of what we considered a good
-reproduction of the target pattern.
+%The threshold value of 0.95 was determined empirically, as signals $r$ with
+%$f(z, z') \ge 0.95$ met our assessment of what we considered a good
+%reproduction of the target pattern.
 
@@
 gradient descend or simulated annealing. But as the computational cost for
-even one learning process is considerable high, these methods were not
+even one learning process is considerably high, these methods were not
 feasible.
@@
 
-Therefore, we optimized $F_C(G)$ using a genetic algorithm, which is inherently
+Therefore, we optimized $F_C$ using a genetic algorithm, which is inherently
 capable of coping with the stochastic fitness function. In each iteration
-of the algorithm, a generation $\mathcal{G}_t = \left\{ G_i~|~1\le i\le N\right\}$
-of $N$ individuals (Generators) is converted into a new one, $\mathcal{G}_{t+1}$,
+of the algorithm, a generation $\mathcal{G}_t = \left\{ G_i~|~1\le i\le M\right\}$
+of $M$ individuals (Genotypes) is converted into a new one, $\mathcal{G}_{t+1}$,
 according to the following scheme:
@@
      $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
- \item[Selection] We choose $p=50\%$ of the most successful members of $G_t$ as survivors.
+ \item[Selection] We choose $50\%$ of the most successful members of $G_t$ as survivors.
      With a small chance, a weaker generator can survive, in order to keep the gene pool more diverse.
-
 \item[Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
@@
 
-\paragraph{Choice of Parameters and Algorithmic Quantities}
-The mutation rate of the parameters $\Theta$ of the genotypes\dots
+%\paragraph{Choice of Parameters and Algorithmic Quantities}
+%The mutation rate of the parameters $\Theta$ of the genotypes\dots
 
@@
 scale (compared to the period to be learned) may disturb the learning.
-That the single genotypes remain quite scattered throughout the whole
-optimization process, though, implies that the exact gain value is not
-critical for the success of a genotype.
+%That the single genotypes remain quite scattered throughout the whole
+%optimization process, though, implies that the exact gain value is not
+%critical for the success of a genotype.
 
@@
 
-But how well did the fitness actually improve during the optimization?
+But how well was the fitness actually improved during the optimization?
 When looking at FIG. \ref{fig:freq_optimization} (d), one sees that the
@@
 
+\begin{figure}[hbct]
+    {\footnotesize\input{pictures/top_optimization}}
+    \caption{
+        Genetic optimization process of genotypes with Erdös-Renyi topology
+        ($N=100$ fixed) for the challenges $C\ix{high}$ and $C\ix{low}$.
+        In each figure, one can see the $50$ genotypes of every Generation
+        (single dots) together with the respective mean values (thick
+        lines). Figures (a) to (c) depict the evolution of the three
+        parameters ($\gain$, $\feedback$, and $\ERprob$) being optimized,
+        while figure (d) shows the development of the fitness values.
+    }
+    \label{fig:freq_optimization}
+\end{figure}
+
 \section{Discussion}
-- Remarks
-    * Influence of initial conditions
-        -> Degenerated initial population poses major problems (does not
-        find same / takes very long to find same parameters as when started
-        with diversity).
-        -> With broad diversity: often fast convergence -> Gen. Optimizer
-        good as optimizing filter, but has hard times to explore new
-        parameter regions if no right genotypes are present (especially isolated
-        maxima of $F_C$). Solution: Could introduce migrants, that supply
-        the gene-pool with underrepresented genotypes.
-    * According to experience: maybe not ideal for finding single set of
-      parameters but rather for successful parameter-ranges (branches?)
+%- Remarks
+    %* Influence of initial conditions
+        %-> Degenerated initial population poses major problems (does not
+        %find same / takes very long to find same parameters as when started
+        %with diversity).
+        %-> With broad diversity: often fast convergence -> Gen. Optimizer
+        %good as optimizing filter, but has hard times to explore new
+        %parameter regions if no right genotypes are present (especially isolated
+        %maxima of $F_C$). Solution: Could introduce migrants, that supply
+        %the gene-pool with underrepresented genotypes.
+    %* According to experience: maybe not ideal for finding single set of
+      %parameters but rather for successful parameter-ranges (branches?)
 
-- Problems
-    * Only few and uncomprehensive number of test / experiments conducted due to the
-      computational cost.
-    * Influence of statistics: Learning success depends stronger on choice
-      of challenges than parameters (partially) -> must be very careful how to
-      interpret the results / should have used fixed array of challenges instead
-      of random sampling
+%- Problems
+    %* Only few and uncomprehensive number of test / experiments conducted due to the
+      %computational cost.
+    %* Influence of statistics: Learning success depends stronger on choice
+      %of challenges than parameters (partially) -> must be very careful how to
+      %interpret the results / should have used fixed array of challenges instead
+      %of random sampling
 

### 2015-08-07 paper/paper.tex
small improvement

@@
 strongly determines whether FORCE learning is able to learn that curve. Still, 
-a good choice of network parameters is still able to improve the success rate for
-certain frequency ranges. Some parameters ($p$, $feedback$) can vary in a relatively large interval
-without influencing the success very much, whereas others are confined to a relatively small interval (gain).
+a good choice of network parameters is able to improve the success rate for
+certain frequency ranges. Some parameters ($\ERprob$, $\feedback$) can vary in a relatively large interval
+without influencing the success very much, whereas others are confined to a relatively small interval ($\gain$).
 
@@
 One problem with the approach presented here is that, in order to use the results of the genetic algorithm, 
-we determined a single set of parameters by taking the mean of the values of the most successful generators, 
-any information about an interdependence between the parameters is lost. We checked that there are no strong
+we determined a single set of parameters by taking the mean of the values of the most successful generators.
+Thereby, any information about an interdependence between the parameters is lost. We checked that there are no strong
 correlations between any pair of parameters by doing 3D plots that showed the two parameters for each generator 

### 2015-08-07 paper/paper.tex
backup commit
updated paper.tex

@@
 
-\begin{figure}[hbct]
+\begin{figure}
     {\footnotesize\input{pictures/freq_optimization}}
@@
         of the three genotypes was tested against constant challenges
-        $C_\omega$ that yielded simple sinusoidal waves with frequency
-        $\omega$ for every seed value. The topology of all genotypes was
+        $C_\nu$ that yielded simple sinusoidal waves with frequency
+        $\nu$ for every seed value. The topology of all genotypes was
         a mere Erdös-Renyi topology.
@@
 
-\begin{figure}[hbct]
+\begin{figure}[]
     {\footnotesize\input{pictures/top_optimization}}

### 2015-08-07 paper/paper.tex
small changes

--- 2015-08-07 paper/paper.tex
@@
 
-Our concrete goal for this article was to examine which choice of genotype
+Our concrete goal for this article was to examine which choices of genotype
 parameters are optimal for learning sinusoidal waves with relatively high
@@
 Based on preliminary simulation results we chose
-$\nu\in[]$ as frequency range for the challenge $C\ix{high}$, producing
-high frequency waves, and $\nu\in[]$ for $C\ix{low}$, producing low
+$\nu\in[0.3, 3]$ as frequency range for the challenge $C\ix{high}$, producing
+high frequency waves, and $\nu\in[0.004, 0.04]$ for $C\ix{low}$, producing low
 frequency waves. 
@@
 A measure that is time-shift independent and maximized for functions of
-identical shape, on the contrary, is the maximum of the cross correlation 
+identical shape is the maximum of the cross correlation 
 \begin{align}

### 2015-08-07 paper/paper.tex
small changes

@@
 performed an optimization in which the networks were an interpolation
-between sparse random connections (Erdös-Renyi) and a ring topology where
+between sparse random connections and a ring topology where
 the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
@@
 Since we are interested in the influence of the topology, the parameters
-gain and feedback are kept constant during the optimization process,
+$\gain$ and $\feedback$ are kept constant during the optimization process,
 leaving the connection probability for ER topology $\ERprob$, the ring

### 2015-08-07 paper/paper.tex
paper

--- 2015-08-07 paper/paper.tex
@@
 
-Our concrete goal for this article was to examine which choices of genotype
+Our concrete goal for this article was to examine which choice of genotype
 parameters are optimal for learning sinusoidal waves with relatively high
@@
 Based on preliminary simulation results we chose
-$\nu\in[0.3, 3]$ as frequency range for the challenge $C\ix{high}$, producing
-high frequency waves, and $\nu\in[0.004, 0.04]$ for $C\ix{low}$, producing low
+$\nu\in[]$ as frequency range for the challenge $C\ix{high}$, producing
+high frequency waves, and $\nu\in[]$ for $C\ix{low}$, producing low
 frequency waves. 
@@
 A measure that is time-shift independent and maximized for functions of
-identical shape is the maximum of the cross correlation 
+identical shape, on the contrary, is the maximum of the cross correlation 
 \begin{align}
@@
 
-Our investigation showed that the frequency of the wave that is to be reconstructed 
-strongly determines whether FORCE learning is able to learn that curve. Still, 
-a good choice of network parameters is able to improve the success rate for
-certain frequency ranges. Some parameters ($\ERprob$, $\feedback$) can vary in a relatively large interval
-without influencing the success very much, whereas others are confined to a relatively small interval ($\gain$).
+%Our investigation showed that the frequency of sinusoidal tasks strongly
+%determines whether FORCE learning is able to learn it or not. Still, a good
+%choice of network parameters, conveying structural information, is able to
+%improve the success rate for certain frequency ranges. 
+%Under certain conditions, some parameters (like $\ERprob$ for low
+%frequencies or $\feedback$ for high frequencies) can vary in a relatively
+%large interval without influencing the success critically.
 
-The code developed for this paper provides a framework which can be used to perform a variety of similar 
-investigations, e.g. the inclusion of feed forward topologies, or the extension to task other than simple 
-sinusoidals. To that end, it would be wise to improve that selection algorithm by using the same tasks
-for all generators in each generation, instead of evaluating them all independently. This would make the 
-selection process more robust because it removes the dependence on the task distribution.
+While we only presented results concerning simple sinusoidals of varying
+frequencies here, the code developed during our project provides
+a framework which can be used to perform a variety of similar and extended
+investigations; for example the inclusion of feed forward topologies and
+arbitrary (higher dimensional) tasks with optional external input pathways
+(e.g.\ as a timer signal to make the network synchronize better).
 
-One problem with the approach presented here is that, in order to use the results of the genetic algorithm, 
-we determined a single set of parameters by taking the mean of the values of the most successful generators.
-Thereby, any information about an interdependence between the parameters is lost. We checked that there are no strong
-correlations between any pair of parameters by doing 3D plots that showed the two parameters for each generator 
-in development with the generations and could not detect any obvious correlation, but have not investigated 
-this problem thoroughly.
+One peculiarity to be recognized throughout our project work is that the
+optimization algorithm does not excel at providing strictly converging,
+single most optimal parameter values, but that we are rather given more or
+less broad ``survival ranges'' for each parameter as result. And though the
+final genotypes varied quite a bit in their fitness, we did not observe
+strong correlations between the different parameters in these ranges when
+doing sporadic tests.
 
-The biggest problem with the current implementation, however, is the large amount of "magic" numbers, 
-i.e. hyper-parameters of the optimization, that are used in the code. Their values were determined
-by an informal search of testing a few optimizations. However, this should mostly affect the performance
-of the optimization algorithm and not the results themselves.
+Indeed, looking at all of our simulations in their entirety, the genotype
+parameters showed less influence on the learning results than we initially
+expected, and during our inquiry it often seamed that the statistics of the
+challenges might overshadow the effect of the parameters -- at least when
+looking at more complex tasks than monofrequent waves. A wise improvement
+would therefore be to make the challenges more deterministic, e.g.\ to use
+the same tasks for all genotypes in each generation instead of drawing them
+independently from some (wide range) distribution.
+
+Another issue to be mentioned with the current implementation is the large
+amount of empirical (``magic") numbers, i.e.\ hyper-parameters of the
+optimization used in the code. Examples hereof are the strength of the
+mutation process or the mechanics of the selection process (actual number
+of survivors, survival chances for weaker genotypes, etc.).  Their values
+were determined by an informal search of testing a few optimizations but
+were not investigated thoroughly. This should, however, mostly affect the
+performance of the optimization algorithm and not the results themselves.
+
+%One problem with the approach presented here is that, in order to use the
+%results of the genetic algorithm, we determined a single set of parameters
+%by taking the mean of the values of the most successful generators.
+%Thereby, any information about an interdependence between the parameters
+%is lost. We checked that there are no strong correlations between any pair
+%of parameters by doing 3D plots that showed the two parameters for each
+%generator in development with the generations and could not detect any
+%obvious correlation, but have not investigated this problem thoroughly.
+
+Nevertheless *ABSCHLUSSSATZ*
+
+%This would certainly make the selection process more robust as it removes
+%the dependence of the task distributions.
+%In general, the genotypes parameters showed less influence than we expected.
+%, whereas others are
+%restricted to more concise regions (???).
+%Some parameters
+%($\ERprob$, $\feedback$) can vary in a relatively large interval without
+%influencing the success very much, whereas others are confined to
+%a relatively small interval ($\gain$).
+
+%To that end, it would be wise to improve that selection
+%algorithm by using the same tasks
+%for all generators in each generation, instead of evaluating them all independently. This would make the 
+%selection process more robust because it removes the dependence on the task distribution.
+
+%One problem with the approach presented here is that, in order to use the results of the genetic algorithm, 
+%we determined a single set of parameters by taking the mean of the values of the most successful generators.
+%Thereby, any information about an interdependence between the parameters is lost. We checked that there are no strong
+%correlations between any pair of parameters by doing 3D plots that showed the two parameters for each generator 
+%in development with the generations and could not detect any obvious correlation, but have not investigated 
+%this problem thoroughly.
 

### 2015-08-07 paper/paper.tex
gain, feedback values

--- 2015-08-07 paper/paper.tex
@@
 Since we are interested in the influence of the topology, the parameters
-$\gain$ and $\feedback$ are kept constant during the optimization process,
+$\gain = 1.5$ and $\feedback = 2$ are kept constant during the optimization process,
 leaving the connection probability for ER topology $\ERprob$, the ring

### 2015-08-07 paper/paper.tex
paper.tex

--- 2015-08-07 paper/paper.tex
@@
     \caption{
-        Genetic optimization process of genotypes with Erdös-Renyi topology
-        ($N=100$ fixed) for the challenges $C\ix{high}$ and $C\ix{low}$.
-        In each figure, one can see the $50$ genotypes of every Generation
-        (single dots) together with the respective mean values (thick
-        lines). Figures (a) to (c) depict the evolution of the three
-        parameters ($\gain$, $\feedback$, and $\ERprob$) being optimized,
-        while figure (d) shows the development of the fitness values.
+        Genetic optimization of genotypes with an fraction $q$ of Erdös-Renyi
+        topology and $1-q$ of ring topology for the challenge $C\ix{high}$.
+        This figure is to be understood like FIG. \ref{fig:freq_optimization}.
     }
-    \label{fig:freq_optimization}
+    \label{fig:top_optimization}
 \end{figure}
@@
 
-Nevertheless *ABSCHLUSSSATZ*
+Nevertheless we were able to produce successful optimization results
+and could indeed show that adaption to high frequencies can not only be
+accomplished by an implausible effect, namely the 
+high feedback values, but also by changing the topological structure of the
+network. In this regard it would certainly be interesting to further
+inquire, how genotypes that allow for a more fine grained control over the
+topology behave under the optimization.
 

### 2015-08-07 paper/paper.tex
abstract

@@
     In this article we examine how structural properties of neural
-    networks, like the size or local synapse densities, affect their
+    networks, like the local synapse densities, affect their
     learning success for various tasks. To do so, we look at a rate based
-    model for neural networks and apply the FORCE rule for the learning
-    process. A sophisticated matching algorithm
+    model for neural networks and apply a learning algorithm called FORCE to let the network learn 
+		to reproduce either low or high frequent sinusoidals. A sophisticated matching algorithm
     allows us to quantify the learning success of a network so that the
-    fitness of structural characteristics, expressed by integer or
-    floating-point parameters, can be evaluated. We then use evolutionary
-    optimization methods on these parameters to show that (1) ring
-    topologies are generally more successful than strictly random
-    topologies, (2) *FAT
-    LIST OF INCREDIBLE RESULTS THAT CHANGE THE UNIVERSE*
+    fitness of structural characteristics can be evaluated. We then use evolutionary
+    optimization methods on these parameters, and found that for high frequencies, the
+		$\feedback$ parameter can be increased  to improve learning results, at the cost of strongly 
+		decreasing the ability to memorize low frequencies. Alternatively, using ring topology instead
+		of random connections has a similar effect, but still retains the ability to learn lower frequencies.
 \end{abstract}

### 2015-08-07 paper/paper.tex
paper.tex

--- 2015-08-07 paper/paper.tex
@@
     \newcommand{\RINGk}{k}
-    \newcommand{\ERtoRING}{r}
+    \newcommand{\ERtoRING}{q}
     \newcommand{\FFlayers}{l}
@@
 $\omega\Ix{fb}$. Optimized parameters of type (1) are the
-occupation probability $p$ for sparse random topologies (Erdös-Renyi
-topology), the neighborhood range $k$ for ring topologies, and a ratio
-parameter $q$ used to interpolate between Erdös-Renyi and ring
+occupation probability $\ERprob$ for sparse random topologies (Erdös-Renyi
+topology), the neighborhood range $\RINGk$ for ring topologies, and a ratio
+parameter $\ERtoRING$ used to interpolate between Erdös-Renyi and ring
 topologies. Optimized parameters of type (2) are the feedback strength
@@
     \caption{
-        The covered frequency range for an unoptimized
-        genotype $G_0$ as well as genotypes $G\ix{high}$ and $G\ix{low}$
-        that are optimized for $C\ix{high}$ respectively $C\ix{low}$. Each
-        of the three genotypes was tested against constant challenges
-        $C_\nu$ that yielded simple sinusoidal waves with frequency
-        $\nu$ for every seed value. The topology of all genotypes was
-        a mere Erdös-Renyi topology.
-        The parameters for the unoptimized genotype $G_0$ were taken to be
-        default values from \cite{FORCE} (which also uses Erdös-Renyi
-        topologies), while we chose genotypes $G\ix{high}$ and $G\ix{low}$
-        that did particularly well when explicitly optimizing for high and
-        low frequencies (see FIG. \ref{fig:freq_optimization} for more
-        details).
+        The fitness values of four differently obtained genotypes when
+        tested against deterministic challenges $C_\nu$ that yield simple
+        sinusoidal waves with frequency $\nu$ for every seed value.  Here
+        $G_0$ is an unoptimized genotype with parameters taken from
+        \cite{FORCE}, whereas the genotypes $G\ix{high}$ and $G\ix{low}$
+        did particularly well during the optimization for $C\ix{high}$ and
+        $G\ix{low}$ depicted in FIG.  \ref{fig:freq_optimization}.  While
+        these three genotypes had strict Erdös-Renyi topologies,
+        $G\ix{high,ring}$ is a successful sample of an optimization for
+        $C\ix{high}$ under an interpolation of Erdös-Renyi and ring topology
+        (see FIG. \ref{fig:top_optimization} in the next section).
     }
+
+        %$G\ix{high,ring}$ is
+        %also optimized for $C\ix{high}$.
+
+        %The covered frequency range for an unoptimized
+        %genotype $G_0$, genotypes $G\ix{high}$ and $G\ix{high,ring}$ optimized for $C\ix{high}$ and $G\ix{low}$
+        %that are optimized for $C\ix{high}$ respectively $C\ix{low}$, . Each
+        %of the three genotypes was tested against constant challenges
+        %$C_\nu$ that yielded simple sinusoidal waves with frequency
+        %$\nu$ for every seed value. The topology of all genotypes was
+        %a mere Erdös-Renyi topology.
+        %The parameters for the unoptimized genotype $G_0$ were taken to be
+        %default values from \cite{FORCE} (which also uses Erdös-Renyi
+        %topologies), while we chose genotypes $G\ix{high}$ and $G\ix{low}$
+        %that did particularly well when explicitly optimizing for high and
+        %low frequencies (see FIG. \ref{fig:freq_optimization} for more
+        %details).
+    %}
     \label{fig:freqrange}
@@
 \subsection{Topology Optimizations}
-To find out whether the topology can be used to improve the learning, we
-performed an optimization in which the networks were an interpolation
-between sparse random connections and a ring topology where
-the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
-fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
-Since we are interested in the influence of the topology, the parameters
-$\gain$ and $\feedback$ are kept constant during the optimization process,
-leaving the connection probability for ER topology $\ERprob$, the ring
-connection length $\RINGk$ and the interpolation value $\ERtoRING$ between
-ER and ring as free parameters.
+To find out whether the topology may be used to improve the learning for
+high frequencies without increasing the feedback, we performed an
+optimization in which the network topologies were an interpolation between
+Erdös-Renyi and ring topology.
+%the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
+%fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
+For this simulation, the parameters $\gain = 1.5$ and $\feedback = 2$ are
+kept constant during the optimization process, leaving the connection
+probability for Erdös-Renyi topology $\ERprob$, the ring connection length $\RINGk$
+and the interpolation value $\ERtoRING$ between Erdös-Renyi and ring as
+free parameters.
 
-The results of the optimization are depicted in FIG. \ref{}. Unfortunately,
-the improvement in fitness after the first few generations in extremely
-low. The development of the ration parameter, however, unambiguously shows
-that the selection process prefers generators that consist mostly of ring
-topology with ring parameter $\RINGk \approx 1$. The ER connection probability
-shows a very broad distribution, which is simply due to the low dependence
-of the generated networks on $\ERprob$ when the usage of ER topology
-$\ERtoRING$ approaches zero.
+The results of the optimization are depicted in FIG.
+\ref{fig:top_optimization}. Unfortunately, the improvement in fitness after the
+first few generations is extremely low. The development of the ration
+parameter, however, unambiguously shows that the selection process prefers
+generators with a high fraction of ring topology and with $\RINGk
+\approx 1.5$. The Erdös-Renyi connection probability $\ERprob$ shows a very
+broad distribution, which is due to the weak dependence of the generated
+networks on $\ERprob$ when $\ERtoRING$ approaches zero.
+
+The fitness curve of $G\ix{high,ring}$ in figure \ref{fig:freqrange}
+confirms, that though $G\ix{high}$ is still better for high frequencies due
+to the strong feedback, the ring topology provides an improvement compared
+to the Erdös-Renyi topology when optimizing for $C\ix{high}$.
 

### 2015-08-07 paper/paper.tex
paper nearly final?

@@
 of the algorithm, a generation $\mathcal{G}_t = \left\{ G_i~|~1\le i\le M\right\}$
-of $M$ individuals (Genotypes) is converted into a new one, $\mathcal{G}_{t+1}$,
+of $M$ individuals (genotypes) is converted into a new one, $\mathcal{G}_{t+1}$,
 according to the following scheme:
-\begin{description}[itemsep=1ex, leftmargin=0pt]
- \item[Evaluation] The fitness $F_i = F_C(G_i)$ of each individual is
+\begin{description}[itemsep=0ex, leftmargin=0pt, format=\normalfont]
+ \item[1. Evaluation] The fitness $F_i = F_C(G_i)$ of each individual is
      measured using at least 50 samples (and adaptively more if the
      $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
- \item[Selection] We choose $50\%$ of the most successful members of $G_t$ as survivors.
+ \item[2. Selection] We choose $50\%$ of the most successful members of $G_t$ as survivors.
      With a small chance, a weaker generator can survive, in order to keep the gene pool more diverse.
-\item[Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
+\item[3. Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
     it into the next generation by either \emph{mutating} or \emph{recombining} the 

### 2015-08-07 paper/paper.tex
second final version

@@
 \author{T. Staudt, E. Schultheis}
-\email{thomas.staudt@stud.uni-goettingen.de, erik.schultheis@uni-goettingen.de}
+\email{thomas.staudt@stud.uni-goettingen.de, erik.schultheis@stud.uni-goettingen.de}
 \affiliation{University of Göttingen}
 
-\date{today}
+\date{\today}
 
 \begin{abstract}
-    In this article we examine how structural properties of neural
-    networks, like the local synapse densities, affect their
-    learning success for various tasks. To do so, we look at a rate based
-    model for neural networks and apply a learning algorithm called FORCE to let the network learn 
-		to reproduce either low or high frequent sinusoidals. A sophisticated matching algorithm
-    allows us to quantify the learning success of a network so that the
-    fitness of structural characteristics can be evaluated. We then use evolutionary
-    optimization methods on these parameters, and found that for high frequencies, the
-		$\feedback$ parameter can be increased  to improve learning results, at the cost of strongly 
-		decreasing the ability to memorize low frequencies. Alternatively, using ring topology instead
-		of random connections has a similar effect, but still retains the ability to learn lower frequencies.
+In this article we examine how structural properties of neural networks,
+like the local synapse density, affect their learning success for various
+tasks. To do so, we look at a rate based model for neural networks and
+apply a learning algorithm called FORCE to teach the network to
+reproduce either low or high frequent sinusoidals. A sophisticated matching
+algorithm allows us to quantify the learning success of the networks so that
+the fitness of structural characteristics (expressed via parameters) can be
+evaluated. We then use evolutionary optimization methods on these
+parameters, 
+and find that the feedback pathway may be strengthened in order
+to improve the learning results for high frequencies but at the cost of
+decreasing the ability to memorize low frequencies.
+%and find that for high frequencies, the feedback pathway
+%can be strengthened to improve learning results at the cost of decreasing
+%the ability to memorize low frequencies. 
+Alternatively, using a ring topology model instead of random connections
+has a similar effect, yet still retains the ability to learn lower
+frequencies better.
 \end{abstract}
@@
 \begin{equation} 
-z = \sum_{j=1}^{N}\omega\Ix{read}_{j}\,r_j
+z = \sum_{j=1}^{N}\omega\Ix{read}_{j}\,r_j~,
 \end{equation}
@@
 pathway, the dynamics $t \mapsto x_i(t)$ of the single neurons are governed
-by \footnote{~Note that we did not include external input signals in this
-formula, as is done in \citep{FORCE} and \citep{RM}.}
+by \footnote{Note that we did not include external input signals in this
+formula, as is done in e.g.\ \citep{RM}.}
 \begin{equation}
@@
 $|z(t) - z'(t)|$ small from the beginning on.
-
-While the authors of \citep{FORCE} update the weights $\lambda\Ix{read}$ in
-fixed intervals, we used an adaptive mechanism to determine how often
+While the authors of \citep{FORCE} update the weights $\omega\Ix{read}$ in
+fixed intervals, we use an adaptive mechanism to determine how often
 learning should occur. 
@@
 In order to find out which structural elements of networks are important
-for learning sucess, we introduce \emph{network genotypes}.
+for learning success, we introduce \emph{network genotypes}.
 %Since we are not interested in single network
@@
 $\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
-$\omega\Ix{fb}$. Optimized parameters of type (1) are the
+$\omega\Ix{fb}$. Optimizable parameters of type (1) are the
 occupation probability $\ERprob$ for sparse random topologies (Erdös-Renyi
-topology), the neighborhood range $\RINGk$ for ring topologies, and a ratio
-parameter $\ERtoRING$ used to interpolate between Erdös-Renyi and ring
-topologies. Optimized parameters of type (2) are the feedback strength
-$\feedback\propto \omega\Ix{fb}$ and the gain value
+topology), the neighborhood range $\RINGk$ for ring topologies (where
+either $\lfloor k\rfloor$ or, with probability $k - \lfloor k\rfloor$,
+$\lfloor k\rfloor+1$ nearest neighbors are connected),
+%fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
+and an interpolation parameter $\ERtoRING$ used to interpolate between Erdös-Renyi
+and ring topology. Optimizable parameters of type (2) are the feedback
+strength $\feedback\propto \omega\Ix{fb}$ and the gain value
 $\gain\propto\omega\Ix{rec}$.
 
-\paragraph{Fitness} The next step was to define the \emph{fitness} of
+\paragraph{Fitness} The next step is to define the \emph{fitness} of
 a given genotype when confronted with a range of different tasks, e.g.
@@
 or low frequencies. 
-Based on preliminary simulation results we chose
-$\nu\in[0.3, 3]$ as frequency range for the challenge $C\ix{high}$, producing
-high frequency waves, and $\nu\in[0.004, 0.04]$ for $C\ix{low}$, producing low
-frequency waves. 
-Optimizing the fitness functions $F_{C\ix{high}}$ or $F_{C\ix{low}}$ thus
-results in genotypes with the desired frequency properties.
+Based on preliminary simulation results we chose $\nu\in[0.3, 3]$ as
+frequency range for the challenge $C\ix{high}$, producing high frequency
+tasks, and $\nu\in[0.004, 0.04]$ for $C\ix{low}$, producing low frequency
+tasks. 
+Optimizing the fitness functions $F_{C\ix{high}}$ or $F_{C\ix{low}}$ will
+thus yield genotypes with the desired frequency properties.
 
@@
 whether a network succeeded to solve a task or not. To this end,
-a measure $f(z', z)$ for the concordance of the target function $t\to z'(t)$ and
+a measure $f(z, z')$ for the concordance of the target function $t\to z'(t)$ and
 the network readout $t\to z(t)$ is sought. 
@@
 
-A measure that is time-shift independent and maximized for functions of
+A measure that is timeshift independent and maximized for functions of
 identical shape is the maximum of the cross correlation 
@@
 Using this measure of quality, the success function for a network
-$\mathcal{N}$ with readout $z$ for the task $z'$ was (by empirical means)
+$\mathcal{N}$ with readout $z$ for the task $z'$ is (by empirical means)
 taken to be 
@@
 \paragraph{Genetic Optimization}
-Due to the stochastic nature of a genotype $G$'s fitness $F_C(G)$,
+Due to the stochastic nature of a genotype's fitness $F_C(G)$,
 a very high number of samples would be required to numerically obtain sufficiently
@@
 gradient descend or simulated annealing. But as the computational cost for
-even one learning process is considerably high, these methods were not
+even one learning process is considerably high, these methods are not
 feasible.
@@
 
-Therefore, we optimized $F_C$ using a genetic algorithm, which is inherently
+Therefore, we optimize $F_C$ using a genetic algorithm, which is inherently
 capable of coping with the stochastic fitness function. In each iteration
@@
      measured using at least 50 samples (and adaptively more if the 
-     $(F_i)_{1\le i\le N}$ are spaced closely, to at most 100).
+     $(F_i)_{1\le i\le N}$ are tightly spaced, to at most 100).
  \item[2. Selection] We choose $50\%$ of the most successful members of $G_t$ as survivors.
      With a small chance, a weaker generator can survive, in order to keep the gene pool more diverse.
-\item[3. Reproduction] We then replace the $(1-p)|G_t|$ elements that did not make
-    it into the next generation by either \emph{mutating} or \emph{recombining} the 
+\item[3. Reproduction] We then replace the $|G_t|/2$ elements that did not make
+    it for the next generation by either \emph{mutating} or \emph{recombining} the 
     survivors and get $\mathcal{G}_{t+1}$. This means we either change
     a single parameter randomly or take two individuals and randomly choose
-    for each parameter from which parent to take the value.   
+    for every parameter from which parent to take the value from.
 \end{description}
@@
 \section{Results}
-\subsection{Optimizing Genotypes For Frequency Ranges}
-Since we wanted to examine how the ideal genotypes vary for different
-types of problems, we examined how the choice of a frequency range affects
-the optimized genotype parameters.
-To this end we looked at sparse random networks (Erdös-Renyi topology with
-connection probability $\ERprob$) and conducted optimizations of the
-parameters $\gain$, $\feedback$ and $\ERprob$ for both high ($C\ix{high}$)
-and low ($C\ix{low}$) frequency challenges (see FIG.
+\subsection{Optimizing Genotypes For High And Low Frequencies}
+Since we want to analyze how the ideal genotypes vary for different types
+of problems, we examine how the choice of either low or high frequencies
+affects the optimized genotype parameters.
+We first look at networks with Erdös-Renyi topology, for which we conduct
+optimizations of the parameters $\gain$, $\feedback$ and $\ERprob$ for both
+high ($C\ix{high}$) and low ($C\ix{low}$) frequency challenges (see FIG.
 \ref{fig:freq_optimization}).
@@
 overshadow the (maybe too slow) internal mechanics for high frequencies,
-while it is the other way round for low frequencies.
+while there is a contrary effect for low frequencies.
 Indeed, FIG. \ref{fig:freq_optimization} (a) shows that genotypes with
@@
 
-But how well was the fitness actually improved during the optimization?
+But how much is the fitness actually improved during the optimization?
 When looking at FIG. \ref{fig:freq_optimization} (d), one sees that the
-fitness for $C\ix{high}$ increased during nearly the whole process, while
-the algorithm failed to further optimize the genotypes' fitness for
+fitness for $C\ix{high}$ increases during nearly the whole process, while
+the algorithm fails to further optimize the genotypes' fitness for
 $C\ix{low}$ after about 20 generations. A major reason for this is
 probably, that the initial population for the $C\ix{low}$ optimization
-already contained genotypes with parameters similar to the (at least
+already contains genotypes with parameters similar to the (at least
 locally) optimal ones.
@@
 indeed succeed at higher frequencies than $G_0$, but that it also fails at
-learning lower frequencies -- which was expected. On the contrary, while
+learning lower frequencies -- which was expected. However, while
 the high frequency capacities of $G\ix{low}$ have decreased when compared
 to $G_0$, the low frequency capacities have only marginally improved. This
-reinforces the impression that the optimization was unable to produce
-genotypes that can cope with low frequencies.
+reinforces the impression that the optimization is unable to produce
+genotypes that can cope with low frequencies significantly better than the
+initial population.
 
@@
 To find out whether the topology may be used to improve the learning for
-high frequencies without increasing the feedback, we performed an
-optimization in which the network topologies were an interpolation between
+high frequencies without increasing the feedback, we perform an
+optimization in which the network topology is an interpolation between
 Erdös-Renyi and ring topology.
@@
 kept constant during the optimization process, leaving the connection
-probability for Erdös-Renyi topology $\ERprob$, the ring connection length $\RINGk$
-and the interpolation value $\ERtoRING$ between Erdös-Renyi and ring as
+probability $\ERprob$ for the Erdös-Renyi topology, the ring neighborhood range $\RINGk$
+and the interpolation value $\ERtoRING$ between Erdös-Renyi and ring topology as
 free parameters.
@@
 \ref{fig:top_optimization}. Unfortunately, the improvement in fitness after the
-first few generations is extremely low. The development of the ration
-parameter, however, unambiguously shows that the selection process prefers
+first few generations is extremely low. The development of the interpolation
+parameter $q$, however, unambiguously shows that the selection process prefers
 generators with a high fraction of ring topology and with $\RINGk
@@
 While we only presented results concerning simple sinusoidals of varying
-frequencies here, the code developed during our project provides
+frequencies in this article, the code developed during our project provides
 a framework which can be used to perform a variety of similar and extended
@@
 less broad ``survival ranges'' for each parameter as result. And though the
-final genotypes varied quite a bit in their fitness, we did not observe
+final genotypes vary quite a bit in their fitness, we did not observe
 strong correlations between the different parameters in these ranges when
@@
 Indeed, looking at all of our simulations in their entirety, the genotype
-parameters showed less influence on the learning results than we initially
+parameters show less influence on the learning results than we initially
 expected, and during our inquiry it often seamed that the statistics of the
@@
 network. In this regard it would certainly be interesting to further
-inquire, how genotypes that allow for a more fine grained control over the
+inquire how genotypes that allow for a more fine grained control over the
 topology behave under the optimization.

### 2015-08-07 paper/paper.tex
final final

@@
     a single parameter randomly or take two individuals and randomly choose
-    for every parameter from which parent to take the value from.
+    for every parameter from which parent to take the value.
 \end{description}
@@
         ($N=100$ fixed) for the challenges $C\ix{high}$ and $C\ix{low}$.
-        In each figure, one can see the $50$ genotypes of every Generation
+        In each figure, one can see the $50$ genotypes of every generation
         (single dots) together with the respective mean values (thick
@@
 
-The fitness curve of $G\ix{high,ring}$ in figure \ref{fig:freqrange}
+The fitness curve of $G\ix{high,ring}$ in FIG. \ref{fig:freqrange}
 confirms, that though $G\ix{high}$ is still better for high frequencies due
